\documentclass{article}
\usepackage{homework-preamble}
\mkanonthms
\begin{document}
\title{General Topology: Assignment 8}
\author{Franz Miltz (UUN: S1971811)}
\date{17 November 2022}
\maketitle

\begin{claim*}
  The topological space $\cc{0,1}^\omega$ where $\cc{0,1}$ is endowed with the discrete
  topology is compact and Hausdorff.
  \begin{proof}
    Compact due to Tikhonov. Hausdorff because any two distinct elements differ in at
    least one place which we may separate locally which is sufficient to make
    sets in the product disjoint.
  \end{proof}
\end{claim*}

\begin{claim*}
  The powerset $\mathcal P\N$ with the chaotic topology is compact but not Hausdorff.
  \begin{proof}
    Compact because the only open cover is $\cc{\mathcal P\N}$. Not Hausdorff because there
    are more than two elements in $\mathcal P\N$ and the topology is chaotic.
  \end{proof}
\end{claim*}

\begin{claim*}
  $(\N, \leq)$ with the Alexandrov topology is compact but not Hausdorff.
  \begin{proof}
    Both follow because the only open set containing $1\in\N$ is $\N$.
  \end{proof}
\end{claim*}

\begin{claim*}
  $(-\N, \leq)$ with the Alexandrov topology is neither compact nor Hausdorff.
  \begin{proof}
    Not compact because every finite set of opens has a minimum element whose predecessor
    is not covered. Not Hausdorff because any open containing the lesser of two distinct
    points also contains the greater one.
  \end{proof}
\end{claim*}

\begin{claim*}
  $S^5\subseteq\R^6$ with the subspace topology is compact and Hausdorff.
  \begin{proof}
    Compact because $S^5$ is a closed and bounded subset of Euclidean space. Hausdorff
    because it is a metric subspace.
  \end{proof}
\end{claim*}

\begin{claim*}
  $\mathbb P^5_\R$ is compact and Hausdorff.
  \begin{proof}
    Compact because $\mathbb P^5_\R$ is homemorphic to a quotient of the compact space $S^5$.
    Hausdorff because because $\mathbb P^5_\R$ is a quotient of the Hausdorff space $\R^6\setminus\cc{0}$
    by a closed equivalence relation.
  \end{proof}
\end{claim*}

\begin{claim*}
  $\hat\Z$ is compact and Hausdorff.
  \begin{proof}
    See $\cc{0,1}^\omega$.
  \end{proof}
\end{claim*}

\begin{claim*}
  $GL\rr{5,\R}\subseteq\R^{25}$ is Hausdorff but not compact.
  \begin{proof}
    Both because it is an unbounded subset of Euclidean space.
  \end{proof}
\end{claim*}

\end{document}
