\documentclass{article}
\usepackage{homework-preamble}
\mkanonthms
\begin{document}
\title{Metric Spaces: Assignment 3}
\author{Franz Miltz (UNN: S1971811)}
\date{7 March 2022}
\maketitle

Let $(X,d_X)$ and $(Y,d_Y)$ be metric spaces.

\begin{claim*}[1a]
   If $f:X\to Y$ is an isometry then $f$ is injective and continuous.
   \begin{proof}
      Let $x,x'\in X$ such that $f(x)=f(x')$. Then $d_Y(f(x),f(x'))=0$ since $d_Y$ is a metric.
      Thus $d_X(x,x')=0$ as $f$ is an isometry. Now, because $d_X$ is a metric, $x=x'$. This proves 
      that $f$ is injective.

      Let $x\in X$ and $\e>0$. Choose $\delta=\e$. Then, for all $x'\in X$ with $d_X(x,x')<\delta$, 
      \begin{align*}
         d_Y(f(x),f(x')) = d_X(x,x') < \delta = \e.
      \end{align*}
      This proves that $f$ is continuous.
   \end{proof}
\end{claim*}

\begin{claim*}[1b]
   Let $f:X\to Y$ be surjective and an isometry. Then $\inv f:Y\to X$ is an isometry.
   \begin{proof}
      Let $y,y'\in Y$. Then, as $f$ is surjective, there exist $x,x'\in X$ such that 
      $f(x)=y$ and $f(x')=y'$. Further, since $f$ is an isometry, $d_X(x,x')=d_Y(y,y')$.     
      However, by definition of $\inv f$, $x=\inv f(y)$ and $x'=\inv f(y')$. The claim is 
      now immediate.
   \end{proof}
\end{claim*}

\begin{claim*}[2]
   Let $f:\R\to \R$ be given by $x\mapsto 0$. Then, with respect to the standard metric 
   $d$ on $\R$, $f$ is continuous but not an isometry.
   \begin{proof}
      Let $x,x'\in X$ be distinct. Then 
      \begin{align*}
         d(x,x') \not= 0 = \abs{0-0} = d(0, 0) = d(f(x), f(x')).
      \end{align*}
      Thus $f$ is not an isometry.

      Let $x\in X$ and $\e>0$. Then, for all $x'\in X$, $d(f(x),f(x'))=0<\e$.
      Thus $f$ is continuous.
   \end{proof}
\end{claim*}

\begin{claim*}[3]
   Let $(X,d_X)$ be a metric space and $C>0$ such that, for all $x,x'\in X$, $d_X(x,x')\leq C$.
   Then, for each $x\in X$, we define $f_x:X\to\R$ by $t\mapsto d_X(x,t)$. Then each $f_x$ is a 
   bounded and continuous function with respect to the standard metric on $\R$.
   \begin{proof}
      Let $x,x'\in X$. Choose $M=C$. Then $\abs{f_x(x')} = \abs{d_X(x,x')} = d_X(x,x') \leq C$. Thus $f_x$ is bounded. 

      Let $x\in X$ and $\e>0$. Choose $\delta=\epsilon$. Then, for all $x',x''\in X$
      with $d_X(x',x'')<\delta$,
      \begin{align*}
         \abs{f_x(x')-f_x(x'')} = \abs{d_X(x,x')-d_X(x,x'')} \leq d_X(x',x'') < \delta = \e
      \end{align*}
      where we used the reverse triangle inequality. Thus $f_x$ is continuous with respect to the 
      standard metric.
   \end{proof}
\end{claim*}

\begin{claim*}[4]
   Let $(X,d_X)$ be a metric space and $C>0$ such that, for all $x,x'\in X$, $d_X(x,x')\leq C$.
   Let $Y$ be the set of all bounded continuous functions $f:X\to\R$ and define the metric 
   $d_Y:Y\times Y\to\R$ by 
   \begin{align*}
      d_Y(f,g) = \sup\{\abs{f(t)-g(t)}:t\in X\}.
   \end{align*}
   Then, for each $x\in X$,  let $f_x:X\to\R$ be given by $t\mapsto d_X(x,t)$. Then the function 
   $g:X\to Y$ given by $x\mapsto f_x$ is an isometry.
   \begin{proof}
      Let $x,x'\in X$. By definition of $d_Y$ we have 
      \begin{align}
         \label{sup}
         d_Y(f_x,f_{x'}) = \sup\{\abs{f_x(t)-f_{x'}(t)}:t\in X\} 
         = \sup\{\abs{d(x,t)-d(x',t)}:t\in X\}.
      \end{align}    
      We note that, for all $t\in X$,
      \begin{align*}
         \abs{d(x,t)-d(x',t)}\leq d(x,t)
      \end{align*}
      by the reverse triangle inequality. Moreover, if $t=x'$, 
      \begin{align*}
         \abs{d(x,t)-d(x',t)}=d(x,t).
      \end{align*} 
      Therefore $d(x,x')$ is an upper bound of the set in (\ref{sup}) which is 
      attained. We conclude that
      \begin{align*}
         d_Y(f_x,f_{x'}) = d(x,x'),
      \end{align*}
      thereby proving the claim.
   \end{proof}
\end{claim*}

\end{document}