\documentclass{article}
\usepackage{homework-preamble}

\begin{document}
\title{FPM: Hand-in 4}
\author{Franz Miltz (UUN: S1971811)}
\date{22 February 2021}
\maketitle

\section*{Problem 1}

Let $\sum_{n=1}^\infty a_n$ be a series of real numbers.

\setcounter{section}{1}
\begin{claim}
	Let $a_n$ be a sequence such that $a_n\geq 0$ for all $n$. If the
	series $\sum_{n=1}^\infty a_n$ converges, then the series
	$\sum_{n=1}^\infty a_n^4$ converges.
\end{claim}
\begin{proof}
	Assume \begin{align*}
		\forall n.\:a_n\geq 0\hs\text{and}\hs \sum_{n=1}^\infty a_n \in \R.
	\end{align*}
	Then by the \emph{Divergence Test} we have
	\begin{align*}
		\lim_{n\to\infty} a_n = 0.
	\end{align*}
	Further, by the \emph{Definition of the Limit} with $\e = 1$
	we know that there exists an $N\in\N$ such that
	\begin{align*}
		n > N\hs\Rightarrow\hs a_n < 1.
	\end{align*}
	Consider the series
	\begin{align*}
		\sum_{n=N+1}^\infty a_n \hs\text{and}\hs \sum_{n=N+1}^\infty a_n^4.
	\end{align*}
	Observe that since $n>N$ we have $a_n<1$ and thus $a_n^4 < a_n$ for all
	$n$. By the \emph{Comparison Test} we conclude that because
	$\sum_{n=N+1}^\infty a_n$ converges, so does $\sum_{n=N+1}^\infty a_n^4$.
	Now consider the partial sums
	\begin{align*}
		S_n = \sum_{k=1}^n a_k^4 \hs\text{and}\hs S'_n = \sum_{k=N+1}^n a_k^4.
	\end{align*}
	Note that we can rewrite $S_n$ as
	\begin{align*}
		S_n = \begin{cases}
			\sum_{k=1}^n a_k^4          & \text{if }n \leq N, \\
			\sum_{k=1}^{N} a_k^4 + S'_n & \text{if }n>N.
		\end{cases}
	\end{align*}
	So we find
	\begin{align*}
		\sum_{n=1}^\infty a_n^4 = \lim_{n\to\infty} S_n
		= \sum_{n=1}^N a_n^4 + \lim_{n\to\infty} S'_n.
	\end{align*}
	Since $\sum_{n=N+1}^\infty a_n^4$ converges and $\sum_{n=1}^N a_n^4$ is
	a real constant, this limit exists and is real. Thus $\sum_{n=1}^\infty a_n^4$ converges.
\end{proof}

\begin{claim}
	Let $a_n$ be a sequence such that $a_n\geq 0$. Then if the series
	$\sum_{n=1}^\infty a_n^4$ converges, $\sum_{n=1}^\infty a_n$ may
	not converge.
\end{claim}

\begin{proof}
	Consider $a_n = 1/n$. Then we have the converging $p$-series
	\begin{align*}
		\sum_{n=1}^\infty a_n^4 = \sum_{n=1}^\infty \frac{1}{n^4}
	\end{align*}
	but the diverging harmonic series
	\begin{align*}
		\sum_{n=1}^\infty a_n = \sum_{n=1}^\infty \frac{1}{n}.
	\end{align*}
\end{proof}

\begin{claim}
	Let $\sum_{n=1}^\infty a_n$ be a converging series. Then
	$\sum_{n=1}^\infty a_n^4$ may not converge.
\end{claim}

\begin{proof}
	Consider the series
	\begin{align*}
		\sum_{n=1}^\infty a_n = \sum_{n=1}^\infty \frac{(-1)^n}{\sqrt[4]{n}}.
	\end{align*}
	Observe that $n^{-1/4}$ is strictly decreasing and thus the series converges
	by \emph{Theorem 3.5.1} from the notes. Now we have
	\begin{align*}
		\sum_{n=1}^\infty a_n^4 = \sum_{n=1}^\infty \frac{1}{n}
	\end{align*}
	which is the harmonic series and diverges.
\end{proof}

\subsection*{Problem 2}

Let $(x_n)_{n\in\N}$ be a sequence of positive real
numbers that converges to a real number $L$.

\setcounter{section}{2}
\setcounter{claim}{0}
\begin{claim}
	If $L>0$, then the series
	\begin{align*}
		\sum_{n=1}^\infty \frac{n^2}{n^4x_n+1}
	\end{align*}
	converges.
\end{claim}
\begin{proof}
	Observe
	\begin{align*}
		\frac{n^2}{n^4x_n+1}\leq\frac{n^2}{n^4x_n} =\frac{1}{n^2x_n}.
	\end{align*}
	Then by the \emph{Definition of the Limit} for any $\e>0$
	we have some $N\in\N$ such that $x_n>L-\e$ for all $n>N$.
	We pick an $\e$ such that $L-\e<1$ and choose $N$ accordingly.
	Thus
	\begin{align*}
		\frac{n^2}{n^4x_n+1}\leq\frac{1}{n^2x_n}\leq \frac{1}{n^2(L-\e)}
	\end{align*}
	for all $n>N$. Since the $p$-series
	\begin{align*}
		\sum_{n=1}^\infty \frac{1}{n^2}
	\end{align*}
	converges, it follows from \emph{Theorem 3.1.12} that
	\begin{align*}
		\sum_{n=1}^\infty \frac{1}{n^2(L-\e)} = \frac{1}{L-\e}\sum_{n=1}^\infty \frac{1}{n^2}
	\end{align*}
	converges and using the \emph{Comparison Test} we find that the series
	\begin{align*}
		\sum_{n=N+1}^\infty \frac{n^2}{n^4x_n+1}
	\end{align*}
	converges as well. By the same reasoning applied to prove \emph{Claim 1.1}
	we conclude that the entire series
	\begin{align*}
		\sum_{n=1}^\infty \frac{n^2}{n^4x_n+1}
	\end{align*}
	converges.
\end{proof}

\begin{claim}
	If $L=0$, then the series
	\begin{align*}
		\sum_{n=1}^\infty \frac{n^2}{n^4x_n+1}
	\end{align*}
	may not converge.
\end{claim}
\begin{proof}
	Consider the sequence
	\begin{align*}
		x_n = \frac{1}{n^4}.
	\end{align*}
	Observe that $x_n>0$ for all $n$ and $x_n\to 0$ as $n\to \infty$.
	We then have the series
	\begin{align*}
		\sum_{n=1}^\infty \frac{n^2}{n^4x_n+1}
		= \frac{1}{2}\sum_{n=1}^\infty n^2.
	\end{align*}
	Since $n^2\to\infty$ as $n\to\infty$, it follows from
	the \emph{Divergence Test} and \emph{Theorem 3.1.12} that the series
	\begin{align*}
		\sum_{n=1}^\infty \frac{n^2}{n^4x_n+1}
	\end{align*}
	diverges.
\end{proof}
\end{document}