\documentclass{article}
\usepackage{homework-preamble}
\usepackage{enumitem}
\mkthms

\title{Introduction to Theoretical Computer Science: Coursework 1}
\begin{document}

\section{Regular languages}

\begin{claim}
	Let $L$ be a regular language. Then its complement $\bar L$ is regular.
\end{claim}
\begin{proof}
	$L$ is regular, thus there exists a DFA, $M(L)=(Q,\Sigma,\delta,q_0,F)$,
	that recognises $L$. We construct another DFA,
	$M(\bar L)=(Q,\Sigma,\delta,q_0,Q\setminus F)$, where a state $q\in Q$ is
	accepting if and only if and only if it is not an accept state in
	$M(L)$. Therefore a word $w\in\Sigma^*$ is accepted by $M(\bar L)$
	if and only if it is not accepted by $M(L)$. Thereby $M(\bar L)$
	recognises $\bar L$.
\end{proof}

\begin{claim}
	Let $L_1,L_2$ be regular. Then the intersection
	\begin{align*}
		L_1\cap L_2 = \{w : w \in L_1 \text{ and } w \in L_2\}
	\end{align*}
	is regular.
\end{claim}
\begin{proof}
	We have
	\begin{align*}
		L_1 \cap L_2 = \Sigma^*\setminus (L_1\cup L_2) = \lneg{L_1\cup L_2}
	\end{align*}
	Since $L_1$ and $L_2$ are regular know that $L_1\cup L_2$ is regular.
	Thus, by the above claim, $\lneg{L_1\cup L_2}=L_1\cap L_2$ is regular.
\end{proof}

\begin{claim}
	A language is regular if and only if there exists an all-NFA that recognises it.
\end{claim}
\begin{proof}
	Let $L$ be regular. Note that there exists a DFA $M=(Q,\Sigma,\delta,q_0,F)$
	that recognises it. We construct an NFA $M'=(Q,\Sigma,\delta',\{q\},F)$
	where $\delta':Q\times\Sigma \to \mathcal{P}(Q)$ is given by $q \mapsto \{\delta(q)\}$.
	Observe that, for all $q\in Q$ and $x\in\Sigma$, $\abs{\delta'(q,x)}=1$.
	Further, $M'$ recognises the same language $L$ as $M$.
	Thus, for any $w\in L$, $\delta^*(q_0,w)\subset F$.

	Now let $M$ be an all-NFA that recognises the language $L$. Then $M$ is an
	NFA, making $L$ regular.
\end{proof}

\begin{claim}
	The language $L=\{0^n1^m2^{m-n}:n\geq m\geq 0\}$ is not regular.
\end{claim}
\begin{proof}
	By contradiction. Assume $L$ is regular. Then, by the \emph{Pumping Lemma},
	there exists a positive integer $p$ such that any word $w\in L$ for which
	$\abs w \geq p$ may be separated into three parts $w=xyz$ such that
	\begin{enumerate}[label=P\arabic*.]
		\item for each $k \geq 0$, $xy^kz\in L$,
		\item $\abs y > 0$,
		\item $\abs{xy}\leq p$.
	\end{enumerate}
	Since $w=0^n1^m2^{m-n}$ for some $n\geq m\geq 0$, there are three cases:
	\begin{enumerate}
		\item $y$ contains no $2$s, i.e. $y=0^s1^t$ for some $s,t\in\Z_{\geq 0}$.
		      Then $w'=xy^0z=0^{n-s}1^{m-t}2^{n-m}$. However, $(n-s)-(m-t)=(n-m)-(s-t)$.
	\end{enumerate}
\end{proof}

\end{document}