\(\tableofcontents\)

\hypertarget{lecture-12-september-20}{%
\section{Lecture 1/2, September 20}\label{lecture-12-september-20}}

All life is made of `ordinary matter' \(\Rightarrow\) Unusual, just 5\%.

\hypertarget{foundations-of-matter}{%
\subsection{Foundations of Matter}\label{foundations-of-matter}}

Matter consists of nucleons (protons \& neutrons) and electrons.

Mass of a nucleon is \(1.66 * 10 ^-27 kg\).

Size of nucleus is \(\approx 10^-15\)

\hypertarget{isotopes}{%
\subsubsection{Isotopes}\label{isotopes}}

Isotopes are atoms of one element that have a different amount of
neutrons.

Lighter isotopes of carbon may be used to look for life. Life tends to
accumulate lighter C isotopes.

\hypertarget{electron-clouds}{%
\subsubsection{Electron clouds}\label{electron-clouds}}

Atomic radius is \(2-5 * 10 ^ -10 m\). Much bigger than nucleus.

\hypertarget{idea}{%
\subsubsection{Idea}\label{idea}}

The Astrobiological Periodic Table.

\hypertarget{bonding}{%
\subsection{Bonding}\label{bonding}}

\hypertarget{ionic-bonding}{%
\subsubsection{Ionic bonding}\label{ionic-bonding}}

\begin{itemize}
\tightlist
\item
  is the force of attraction between positively and negatively charged
  ions
\item
  most ionic compounds are crystaline solid
\end{itemize}

``Atoms don't like spare electron shells'' \(\Rightarrow\) Lowest energy
state. (e.g.~Sodium + Chlorine)

Atoms arrange themselves in a grid-like structure, naturally.

\textbf{NaF}: - energy bond = \(300\times 10 ^ -21 J\) -
\(E\approx k_b T\) is an atoms energy due to thermal motion.

Ionic bonds hold proteins together.

\hypertarget{covalent-bonding}{%
\subsubsection{Covalent Bonding}\label{covalent-bonding}}

\begin{itemize}
\tightlist
\item
  takes place between atoms with small differences in charge which are
  close to each other in the periodic table or identical atoms
  (e.g.~hydrogen, carbon)
\item
  is formed by sharing an outer electron
\item
  it's a stable bond, 150 times less thermal energy than bonding energy
\item
  prominent in extreme termperatures
\end{itemize}

\(\Rightarrow\) All of this is universal.

\hypertarget{metallic-bonding}{%
\subsubsection{Metallic Bonding}\label{metallic-bonding}}

\begin{itemize}
\tightlist
\item
  is the type of bonding found in metal elements - \textbf{not important
  for life}. THis is the electrostatic force of attraction between
  positively charged ions and declocalized electrons.
\item
  there is no organism with a solid metal in their bodies; why?

  \begin{itemize}
  \tightlist
  \item
    rusting
  \item
    not enough energy
  \end{itemize}
\end{itemize}

\hypertarget{van-der-waals-forces}{%
\subsubsection{van der Waals Forces}\label{van-der-waals-forces}}

\hypertarget{dipole-dipole-forces}{%
\paragraph{Dipole-dipole Forces}\label{dipole-dipole-forces}}

\begin{itemize}
\item
  each molecule has a charge differences accros it's length

  \(\Rightarrow\) small permanent dipole \(\Rightarrow\) two dipoles
  attract each other
\end{itemize}

\hypertarget{induced-dipole}{%
\paragraph{Induced dipole}\label{induced-dipole}}

\begin{itemize}
\tightlist
\item
  a dipole inserts a charge
\end{itemize}

\hypertarget{dispersion-forces}{%
\paragraph{Dispersion Forces}\label{dispersion-forces}}

\begin{itemize}
\tightlist
\item
  exists between all atoms and molecules
\item
  due to uneven distribution of electrons in atom (by probability)
\end{itemize}

\hypertarget{hydrogen-bonding}{%
\subsubsection{Hydrogen Bonding}\label{hydrogen-bonding}}

\begin{itemize}
\tightlist
\item
  found in molecules containing OH groups
\item
  attraction of electrons of other molecules
\item
  \(1/10\) of a covalent bond
\end{itemize}

Hydrogen bonds hold DNA together. Useful because they need to be broken
and reassembled often.

\begin{quote}
Summary of bonding types
\end{quote}

\begin{quote}
What you need to know
\end{quote}

\hypertarget{states-of-matter}{%
\subsection{States of Matter}\label{states-of-matter}}

sorted by increasing energy - solid - liquid - gas - plasma (99\% of all
ordinary matter)

(organisms are mad up of \(0.046\)\% of matter)

triple point: point on a phase diagram where all three states of matter
co-exist

\begin{quote}
supercritical fluid
\end{quote}

e.g.~Mars: - surface is not suited for life - anything below that might
be suitable

\hypertarget{anomaly-of-water}{%
\subsubsection{Anomaly of water}\label{anomaly-of-water}}

\begin{itemize}
\item
  solid water has higher density
\item
  higher pressure means more liquid than solid

  \(\Rightarrow\) Habitat for life under water in winter.
\end{itemize}

\hypertarget{other-states-of-matter}{%
\subsubsection{Other states of matter}\label{other-states-of-matter}}

\hypertarget{plasma}{%
\paragraph{Plasma}\label{plasma}}

Hot plasma in nature: - stars - lightning - Aurora Borealis

Not in structure of life. Yet, it's important for life.

\hypertarget{degenerate-matter}{%
\paragraph{Degenerate Matter}\label{degenerate-matter}}

\begin{itemize}
\tightlist
\item
  collection of free, non-interacting particles
\item
  occurs at very high density (\(>10^3 \frac{kg}{cm^3}\))
\item
  electrons can't fit in the smaller atom area
\end{itemize}

Exists in white dwarf stars. (\(>10^5 \frac{kg}{cm^3}\))

Why is life on neutron stars hard?

\begin{itemize}
\tightlist
\item
  neutron stars tend to leave their star systems
\end{itemize}

\hypertarget{lecture-4-september-23}{%
\section{Lecture 4, September 23}\label{lecture-4-september-23}}

\hypertarget{fundamentals-of-life}{%
\subsection{Fundamentals of Life}\label{fundamentals-of-life}}

\hypertarget{molecules}{%
\subsubsection{Molecules}\label{molecules}}

Six major elements: C, H, N, O, P, S

\begin{itemize}
\tightlist
\item
  Carbon
\item
  Hydrogen
\item
  Nitrogen
\item
  Oxygen
\item
  Phosphorus
\item
  Sulfur
\end{itemize}

\(\Rightarrow\) CHNOPS

Other elements (Calcium, Iron, \ldots) are important as well.

Carbon is the main backbone of all life on earth. \(\Rightarrow\) Carbon
based life.

\hypertarget{why-carbon}{%
\paragraph{Why Carbon?}\label{why-carbon}}

\begin{itemize}
\tightlist
\item
  Bond energy between C and HNOPS are similar \(\Rightarrow\) Bond
  change does not require a lot of energy.
\item
  Links are quite stable.
\item
  Wide range of double and triple bond configurations.
\end{itemize}

\begin{quote}
Capacity to form chains with complex branching and networking of rings.
Millions of possible `organic' compounds can be formed.
\end{quote}

\(\Rightarrow\) Selection of elements is not random.

\hypertarget{alternatives}{%
\paragraph{Alternatives?}\label{alternatives}}

What about Silicon?

\begin{itemize}
\tightlist
\item
  it is possible to make organo-silicon compounds
\item
  complex chained compounds are also possible
\item
  silicon forms stable compounds with oxygen (\texttt{SiO})

  \begin{itemize}
  \tightlist
  \item
    binds to 4 oxygen atoms instead of 2
  \item
    those atoms connect to others
  \item
    VERY stable
  \item
    ROCKS, not that good\ldots{}
  \end{itemize}
\item
  there has been a lot of silicon chemistry, but no life has resulted of
  it
\item
  while there's oxygen, silicon-based life is unlikely
\item
  life on earth DOES use silicon

  \begin{itemize}
  \tightlist
  \item
    e.g.~plants, for stability; spicules in sponges
  \end{itemize}
\end{itemize}

\hypertarget{major-classes-of-molecules}{%
\paragraph{Major classes of
molecules}\label{major-classes-of-molecules}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Proteins
\item
  Sugars (structural support)
\item
  Lipids (membranes and energy)
\item
  Nucleic acids (information storage)
\end{enumerate}

\hypertarget{proteins}{%
\subparagraph{Proteins}\label{proteins}}

\begin{itemize}
\tightlist
\item
  constructed from amino acids
\item
  structural roles
\item
  built by `condensation' or `dehydration synthesis' resulting in a
  peptide link
\item
  charges in amino acids cause secondary and tertiary structure
\end{itemize}

Amino acids: - are chiral: identical structure, different orientation -
L-amino acids - D-amino acids - based on rotation of light - mirrored
life might be possible

normally: L-amino acids, D-sugar possibly: D-amino acids, L-sugar

\hypertarget{sugars}{%
\subparagraph{Sugars}\label{sugars}}

\begin{itemize}
\tightlist
\item
  used to store energy
\item
  structural support
\item
  build by dehydration reaction
\end{itemize}

\hypertarget{lipids}{%
\subparagraph{Lipids}\label{lipids}}

\begin{itemize}
\tightlist
\item
  long chains
\item
  in membranes
\item
  may store energy
\end{itemize}

\hypertarget{nucleic-acids}{%
\subparagraph{Nucleic acids}\label{nucleic-acids}}

\begin{itemize}
\tightlist
\item
  stores information
\item
  made of ribose sugar, phosphate groub and base pairs
\item
  complicated molecules by simple chemistry
\end{itemize}

\begin{quote}
Meteorites contain amino acids and more\ldots{}
\end{quote}

\hypertarget{life-needs-water}{%
\subsubsection{Life needs water}\label{life-needs-water}}

\begin{itemize}
\tightlist
\item
  medium for chemical reactions
\item
  carries nutrients and energy supplies
\item
  helps maintain thermal balance (self-regulation)
\end{itemize}

Advantages of water: - can dissolve irons - large heat capacity - is
cosmically abundant

\hypertarget{alternatives-1}{%
\paragraph{Alternatives?}\label{alternatives-1}}

Ammonia:

\begin{itemize}
\tightlist
\item
  dissolves many chemicals
\item
  potentially alows for ammonia-based bonds betwee chemicals similar to
  amino acids to make long chains
\item
  less vicious than water
\item
  quite high heat vapourisation
\end{itemize}

Hydrogen fluoride:

\begin{itemize}
\tightlist
\item
  wide temperature range
\item
  dissolves a wide range of substances
\end{itemize}

but - rare - very reactive and tends to destroy organic compounds

\hypertarget{what-you-need-to-know}{%
\subsection{What you need to know}\label{what-you-need-to-know}}

\begin{itemize}
\tightlist
\item
  know the CHNOPS
\item
  why carbon; why not silicon
\item
  classes of biological molecules; uses
\item
  chirality
\item
  structure of DNA, what's RNA
\item
  characteristics of water; alternatives
\end{itemize}

\hypertarget{lecutre-45-september-27}{%
\section{Lecutre (4/5), September 27}\label{lecutre-45-september-27}}

\begin{itemize}
\tightlist
\item
  get your barcode from 4309 (!!!)
\end{itemize}

\hypertarget{fundamentals-of-life-1}{%
\subsection{Fundamentals of Life}\label{fundamentals-of-life-1}}

Where are universal structures in life?

\hypertarget{cells}{%
\subsubsection{Cells}\label{cells}}

All life on earth is made of cells.

Made of - membrane - increases pressure, makes chemistry possible - all
living things have membranes (apart from viruses) - information system -
energy system

\hypertarget{types-of-cells}{%
\paragraph{Types of cells}\label{types-of-cells}}

Prokaryotes (no nucleus) - bacteria and archaea - \(1-10\mu m\)

Eukaryotes (nucleus) - \(10-100\mu m\) - nucleus to encapsule DNA -
organelles that compartmentalise

Most life is prokaryotic. - Anthony van Leeuwenhoek discovered it -
something like \(10^30\) - microbes evolved, today's are different from
the first ones - they have complicated parts

\(\Rightarrow\) We don't have any example of the most simple life!

\hypertarget{membranes}{%
\paragraph{Membranes}\label{membranes}}

Phospholipids!

\begin{itemize}
\tightlist
\item
  Two differently polarized ends.
\item
  Spontaneous arrangement in membrane structure.
\item
  it will form a sphere-like object (minimization of surface tension)
\end{itemize}

This is not difficult to achieve; it can be done with molecules from
Asteroids.

Two kinds: Gram negative and Gram positive.

Membrane needs to allow communcation and transfer of matter.

\hypertarget{peptidoglycan}{%
\paragraph{Peptidoglycan}\label{peptidoglycan}}

A cell protecting `mesh'.

Built out of D-amino acids. (Weird.)

\hypertarget{dna}{%
\paragraph{DNA}\label{dna}}

Deoxyribonucleic acid

\begin{itemize}
\tightlist
\item
  double-helix structure
\item
  encoding through bases
\item
  two pairs: A-T and G-C
\item
  replication through redundancy
\item
  biology goes only from DNA to Proteins; never the other way
\end{itemize}

\hypertarget{dna-vs-rna}{%
\paragraph{DNA vs RNA}\label{dna-vs-rna}}

\begin{longtable}[]{@{}ll@{}}
\toprule
DNA & RNA\tabularnewline
\midrule
\endhead
Double Stranded & Single Stranded\tabularnewline
Deoxyribose sugar & Ribose Sugar\tabularnewline
Bases: A G C T & Bases: A C G U\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{transcription}{%
\paragraph{Transcription}\label{transcription}}

\(DNA \rightarrow mRNA\)

RNA Polymerase crates mRNA.

Function: - Unwind DNA sequence.

\hypertarget{translation}{%
\paragraph{Translation}\label{translation}}

\(mRNA \rightarrow protein\)

\begin{itemize}
\tightlist
\item
  Second stage of protein production
\item
  using ribsomal RNA (rRNA) - very conserved type of RNA
\item
  tRNA goes into Ribosome
\item
  complementary code of amino acid
\item
  3 letters = 1 amino acids
\item
  codon = coding triplet of mRNA bases
\item
  Genetic code = inventory of linkages between nucleotide triplets and
  the amino acids they code for
\item
  A gene is a segment of DNA that is useful.
\item
  there are 64 possible combinations
\item
  yet, only twenty amino acids exist
\item
  redundancy in DNA allows for tolerance of mutation/errors
\end{itemize}

\hypertarget{what-you-need-to-know-1}{%
\subsection{What you need to know}\label{what-you-need-to-know-1}}

\begin{itemize}
\tightlist
\item
  major featurs of cells
\item
  how membranes form and draw a basic membrane
\item
  Understand that there are twoo different types of bacterial membrane
\item
  Knw the difference between RNA and DNA
\item
  Know what a ribosome is
\item
  Understand and be able to explain transcription and translation
\item
  Given a codon table, be able write DNA, RNA, protein sequences
\end{itemize}

\hypertarget{phylogenetic-trees-diversity-of-life}{%
\subsection{Phylogenetic Trees / Diversity of
Life}\label{phylogenetic-trees-diversity-of-life}}

All life on earth is related.

Two options: - A rare event that resulted in us, thus we're special. - A
lot of different forms of life once existed, yet only one became
dominant.

What is LUCA? (Last Universal Common Ancestor) - way simpler than todays
life - scientists tried to recreate it's proteins

\hypertarget{classifying-by-resemblance}{%
\subsubsection{Classifying by
resemblance}\label{classifying-by-resemblance}}

\begin{quote}
``All cats look the same.'' (Aristotle)
\end{quote}

This does not work for microbes.

\hypertarget{terms}{%
\subsubsection{Terms}\label{terms}}

\begin{itemize}
\tightlist
\item
  \textbf{Phylogeny} is the evolutionary history of a species or group
  of related species.
\item
  The discipline of \textbf{systematics} classifies organisms and
  determines their evolutionary relationships.
\item
  \textbf{Taxonomy} is the ordered division and naming of organisms.
\end{itemize}

\hypertarget{system-of-species-classification}{%
\subsubsection{System of species
classification}\label{system-of-species-classification}}

\begin{itemize}
\tightlist
\item
  binomial nomenclature
\item
  first part is the genus (e.g.~``Canis'' = dog)
\item
  the second part, called the specific epithet, is unique for each
  species within the genus (e.g.~``Lupus'' =\textgreater{} Wolf)
\end{itemize}

\hypertarget{hierarichal-classification}{%
\paragraph{Hierarichal
classification}\label{hierarichal-classification}}

\begin{itemize}
\tightlist
\item
  Species
\item
  Genus
\item
  Family
\item
  Order
\item
  Class
\item
  Phylum
\item
  Kingdom
\item
  Domain
\end{itemize}

\hypertarget{phylogenetic-trees}{%
\paragraph{Phylogenetic Trees}\label{phylogenetic-trees}}

\begin{itemize}
\tightlist
\item
  Nodes: branching point
\item
  Outgroup: Species that's very different to cluster other secies, one
  is interested in together
\item
  Root: Last common ancestor of all the taxa
\item
  Sister taxa: Taxa that are direct result of a node
\end{itemize}

Problm with phylogenies: - looks can be deceiving - similar species
might have evolved in different parts of the world independently

\(\Rightarrow\) Conversion evolution / Analogy.

e.g.~Dolphins / Fish.

Homology can be distinguished from analogy by comparing fossil evidence
and the degree of complexity.

How to create trees?

\begin{itemize}
\tightlist
\item
  ribosomes change only in small ways
\item
  thus you can see the differences and similarities at the same time
\end{itemize}

\hypertarget{mutations}{%
\paragraph{Mutations}\label{mutations}}

Over billions of years there will be change. Life cannot avoid the
second law of thermal dynamics.

\begin{itemize}
\tightlist
\item
  Transversion
\item
  Deletion
\item
  Transition
\item
  Insertionk
\end{itemize}

\hypertarget{methods-today}{%
\paragraph{Methods today}\label{methods-today}}

Reverse engineering through observed rate of change with computer
software.

Multiple trees are generated and occurences of certain relations
indicate our confidence in the structure.

\hypertarget{sorting-hormology-from-analogy}{%
\paragraph{Sorting hormology from
analogy}\label{sorting-hormology-from-analogy}}

e.g.~AGA -\textgreater{} AAA -\textgreater{} AGA

Mathematics allow you to compare the rest of the sequence.

\begin{quote}
MATHEMATICS OF PHYLOGENETIC TREES ARE NO LONGER IN EXAMS.
\end{quote}

\hypertarget{evolutionary-hypotheses}{%
\subsubsection{Evolutionary Hypotheses}\label{evolutionary-hypotheses}}

\begin{itemize}
\tightlist
\item
  Type 1 Inference: Both ancestors have trait
\item
  Type 2 Inference: One of the ancestors has trait
\item
  Type 3 Inference: Neither has trait
\end{itemize}

\hypertarget{complicating-the-tree-of-life}{%
\subsubsection{Complicating the tree of
life}\label{complicating-the-tree-of-life}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Microorganisms have encorporated others
\item
  Bacteria can take up DNA from the environment
\item
  DNA can be injected into bacteria (i.e.~viruses)
\item
  Microorganisms can transfer DNA
\end{enumerate}

The tree of life is not as tidy as we'd wish; not as Darwin thought.

\hypertarget{what-you-need-to-know-2}{%
\subsection{What you need to know}\label{what-you-need-to-know-2}}

\begin{itemize}
\tightlist
\item
  principles behind the hierarchical classification
\item
  difference between homology and analogy
\item
  know what a nonophyletic group is
\item
  how DNA is used to build PG trees
\item
  understand horizontal gene transfer; give two examples
\item
  know how to use pholygenetic trees to test hypothesis; example
\item
  hypothetical implications of these ideas for the search of life
  elsewhere
\end{itemize}

\hypertarget{lecture-7-september-30}{%
\section{Lecture 7, September 30}\label{lecture-7-september-30}}

\begin{quote}
Energy and Life
\end{quote}

\hypertarget{definitions}{%
\subsection{Definitions}\label{definitions}}

\begin{itemize}
\tightlist
\item
  \textbf{Catabolism}: The processes by which a living organism obtains
  its energy and raw materials from nutrients.
\item
  \textbf{Anabolism}: The processes by which energy and raw materials
  are used to build macromolecules and cellular structures.
\item
  \textbf{Metabolism}: The processes of catabolism and anabolism.
\end{itemize}

Classification: - Source of energy - Source of carbon

Specifically: - Photogrophs: can use light energy - Chemeotrophs: must
obtain energy from oxidation-reduction of chemical compounds -
Autotrophs: draw carbon from carbon dioxide - Heterotrophs: carbon from
organic compounds - Mixotrophic: carbon can be obtained from both
organic compounds and by fixing carbon dioxide

You may combine those. (e.g.~Chemoautotrophs)

\hypertarget{aerobic-respiration}{%
\subsection{Aerobic respiration}\label{aerobic-respiration}}

Terminal electron acceptor is oxygen.

\(C_6 H_12 O_6 \rightarrow 6 CO_2 + 6 H_2 O\)

\begin{itemize}
\tightlist
\item
  \(C_6 H_12 O_6\): Electron donor
\item
  \(CO_2\): Electron acceptor
\end{itemize}

How do we get the energy from the reaction?

\begin{itemize}
\tightlist
\item
  energy from the free electron is used to move protons (\(H^+\)) out of
  a cell
\item
  creates an osmotic gradient (differing concentrations inside/outside)
\item
  proton moves back through ATP synthase
\item
  \(ADP + P \rightarrow ATP\)
\end{itemize}

\(\Rightarrow\) Energy from the electron is captured in a chemical bond.

\hypertarget{anaerobic-respiration}{%
\subsection{Anaerobic respiration}\label{anaerobic-respiration}}

Terminal electron acceptor is something other than oxygen:

\begin{itemize}
\tightlist
\item
  Nitrate (\(NO_3^-\)) e.g.~Pseudomonas, Bacillus
\item
  Sulfate (\(SO_4^{2-}\)) e.g.~Desulfovibrio
\item
  \(CO_2\)
\end{itemize}

\hypertarget{chemolithotrophy}{%
\subsection{Chemolithotrophy}\label{chemolithotrophy}}

Very interesting because \textbf{donors} are inorganic compounds.

\begin{itemize}
\tightlist
\item
  \(H_2 S\)
\item
  \(H_2\)
\item
  \(NH_4^+\)
\item
  \(Fe^{2+}\)
\end{itemize}
