\documentclass{article}
\usepackage[a4paper]{geometry}
\geometry{tmargin=3cm, bmargin=3cm, lmargin=2cm, rmargin=2cm}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\DeclareMathOperator{\sech}{sech}
\DeclareMathOperator{\csch}{csch}
\DeclareMathOperator{\arccot}{\text{cot}^{-1}}
\DeclareMathOperator{\arccsc}{\text{csc}^{-1}}
\DeclareMathOperator{\arccosh}{\text{cosh}^{-1}}
\DeclareMathOperator{\arcsinh}{\text{sinh}^{-1}}
\DeclareMathOperator{\arctanh}{\text{tanh}^{-1}}
\DeclareMathOperator{\arcsech}{\text{sech}^{-1}}
\DeclareMathOperator{\arccsch}{\text{csch}^{-1}}
\DeclareMathOperator{\arccoth}{\text{coth}^{-1}} 
\DeclareMathOperator{\dom}{dom}
\DeclareMathOperator{\st}{s.t.}
\usepackage{amsthm}
\usepackage{siunitx}
\usepackage{nicefrac}
\newtheoremstyle{sltheorem} {}                % Space above
{}                % Space below
{\upshape}        % Theorem body font % (default is "\upshape")
{}                % Indent amount
{\bfseries}       % Theorem head font % (default is \mdseries)
{.}               % Punctuation after theorem head % default: no punctuation
{ }               % Space after theorem head
{}                % Theorem head spec
\theoremstyle{sltheorem}
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\begin{document}
\tableofcontents
\setcounter{section}{3}
\section{Applications of differentiation}
\setcounter{subsection}{6}
\subsection{Antiderivatives}
\begin{definition}
    A function $F$ is called an \textbf{antiderivative} of $f$ on an interval $I$ if $F'(x)=f(x)$ for all $x$ in $I$.
\end{definition}
\begin{theorem}
    If $F$ is an antiderivative of $f$ on an interval $I$, then the most general antiderivative of $f$ on $I$ is
    \begin{align*}
        F(x) + C
    \end{align*}
    where $C$ is an arbitrary constant.
\end{theorem}
\begin{theorem}
    Particular antiderivatives of common functions:
    \begin{itemize}
        \item $\int cf(x)dx = c\int f(x)dx$
        \item $\int f(x) + g(x) dx = \int f(x)dx + \int g(x)dx$
        \item $\int x^n dx=\frac{x^{n+1}}{n+1}+C$
        \item $\int \frac{1}{x}dx=\ln |x|+C$
        \item $\int e^xdx=e^x+C$
        \item $\int \cos x dx = \sin x+C$
        \item $\int \sin x dx = -\cos x +C $
        \item $\int \sec^2 x dx = \tan x + C$
        \item $\int \sec x \tan x dx = \sec x + C$
        \item $\int \frac{1}{\sqrt{1-x^2}}dx = \sin^{-1} x + C$
        \item $\int \frac{1}{1+x^2}dx = \arctan x + C$
        \item $\int \cosh x dx = \sinh x + C$
        \item $\int \sinh x dx = \cosh x + C$
    \end{itemize}
\end{theorem}
\section{Integrals}
\subsection{Areas and Distances}
\begin{definition}
  The \textbf{area} $A$ of the region $S$ that lies under the graph of the continuous function $f$ is the limit of the sum of the areas of appoximating rectangles:
  \begin{align*}
      A=\lim_{n\to\infty}R_n = \lim_{n\to \infty}\left(f(x_1)\Delta x + f(x_2)\Delta x + \cdots + f(x_n)\delta x\right)
  \end{align*}
\end{definition}
\subsection{The definite integral}
\begin{definition}
    If $f$ is a function defined on $[a,b]$, the \textbf{definite integral of $f$ from $a$ to $b$} is the number
    \begin{align*}
        \int_a^b f(x)dx=\lim_{\max \Delta x_i\to0}\sum_{i=1}^n f(x_i^*)\Delta x_i
    \end{align*}
    provided that this limit exists. If it does exist, we say that $f$ is \textbf{integrable} on $[a,b]$.
\end{definition}
\begin{theorem}
    If $f$ is continuous on $[a,b]$, or if $f$ has only a finite number of jump discontinuities, then $f$ is integrable on $[a,b]$; that is, the definite integral $\int_a^b f(x)dx$ exists.
\end{theorem}
\begin{theorem}
    If $f$ is integrable on $[a,b]$, then \begin{align*}
        \int_a^b f(x)dx = \lim_{n\to\infty}\sum_{i=1}^nf(x_i)\Delta x
    \end{align*}
    where \begin{align*}
    \Delta x = \frac{b-a}{n} \text{ and } x_i=a+i\Delta x
    \end{align*}
\end{theorem}
\begin{theorem}
    \textbf{The midpoint rule}.
    \begin{align*}
        \int_a^bf(x)dx \approx \sum_{i=1}^nf(\bar{x_i})\Delta x = \Delta x (f(\bar{x_i})+\cdots+f(\bar{x_n}))
    \end{align*}
    where
    \begin{align*}
        \Delta x = \frac{b-a}{n},\\
        \bar{x_i} = \frac{1}{2}(x_{i-1}+x_i).
    \end{align*}
\end{theorem}
\begin{theorem}
    \textbf{Properties of integrals}.\\
    Suppose all the following integrals exist and let $c$ be any constant.
    \begin{enumerate}
        \item $\int_a^b cdx = c(b-a)$
        \item $\int_a^b(f(x)+g(x))dx=\int_a^b+\int_a^bg(x)dx$
        \item $\int_a^b(cf(x))dx=c\int_a^bf(x)dx$
        \item $\int_a^b(f(x)-g(x))dx=\int_a^bf(x)dx-\int_a^bg(x)dx$
        \item $\int_a^cf(x)dx+\int_c^bf(x)dx = \int_a^bf(x)dx$
        \item If $f(x)\geq 0$ for $x\in[a,b]$, then $\int_a^b f(x)dx \geq 0$.
        \item If $f(x)\geq g(x)$ for $x\in[a,b]$, then $\int_a^b f(x)dx\geq \int_a^b g(x)dx$.
        \item If $f(x)\in[m,M]$ for $x\in[a,b]$, then 
        \begin{align*}
            m(b-a)\leq \int_a^b f(x)dx \leq M(b-a)
        \end{align*}
    \end{enumerate}
\end{theorem}
\subsection{Evaluating definite integrals}
\begin{theorem}
    \textbf{Net change theorem}. The integral of a rate of change is the net change:
    \begin{align*}
        \int_a^bF'(x)dx=F(b)-F(a)
    \end{align*}
\end{theorem}
\subsection{The fundamental theorem of calculus}
\begin{theorem}
   Suppose $f$ is continuous on $[a,b]$.
   \begin{enumerate}
       \item If $g(x)=\int_a^x f(t)dt$, then $g'(x)=f(x)$.
       \item $\int_a^b f(x)dx=F(b)-F(a)$, where $F$ is any antiderivative of $f$, that is $F'=f$.
   \end{enumerate}
\end{theorem}
\begin{definition}
    The \textbf{average value of $f$} on the interval $[a,b]$ is
    \begin{align*}
        f_{ave}=\frac{1}{b-a}\int_a^b f(x)dx.
    \end{align*}
\end{definition}
\begin{theorem}
    \textbf{The mean value theorem for integrals.} If $f$ is continuous on $[a,b]$, then there exists a number $c$ in $[a,b]$ such that
    \begin{align*}
        f(c)=f_{ave}=\frac{1}{b-a}\int_a^b f(x)dx
    \end{align*}
    that is,
    \begin{align*}
        \int_a^b f(x)dx=f(x)(b-a).
    \end{align*}
\end{theorem}
\subsection{The substitution rule}
\begin{theorem}
    \textbf{The substitution rule.}
    If $u=g(x)$ is a differetiable function whose range is an interval $I$ and $f$ is continuous on $I$, then
    \begin{align*}
        \int f(g(x))g'(x)\:dx=\int f(u)\:du
    \end{align*}
\end{theorem}
\begin{theorem}
    \textbf{The substitution rule for definite integrals.}
    If $g'$ is continuous on $[a,b]$ and $f$ is continuous on the range of $u=g(x)$, then 
    \begin{align*}
        \int_a^b f(g(x))g'(x)\:dx=\int_{g(a)}^{g(b)}f(u)\:du.
    \end{align*}
\end{theorem}
\begin{theorem}
    \textbf{Integrals of symmetric functions.} Suppose $f$ is continuous on $[-a,a]$.
    \begin{enumerate}
        \item If $f$ is even, then $\int_{-a}^a f(x)\:dx= 2\int_0^a f(x)\:dx$.
        \item If $f$ is odd, then $\int_{-a}^a f(x)\:dx = 0$.
    \end{enumerate}
\end{theorem}
\section{Techniques of integration}
\begin{theorem}
    \begin{align*}
        \int u dv = uv - \int v du
    \end{align*}
\end{theorem}
\begin{definition}
    \textbf{of an improper integral of type 1.}
    \begin{enumerate}
        \item If $\int_a^t f(x)\:dx$ exists for every number $t\geq a$, then
        \begin{align*}
            \int_a^\infty f(x)\:dx = \lim_{t\to \infty}\int_a^t f(x)\:dx.
        \end{align*}
        provided this limit exists (as a finite number).
        \item If $\int_t^b f(x)\:dx$ exists for every $t\leq b$, then
        \begin{align*}
            \int_{-\infty}^b f(x)\: dx = \lim_{t\to-\infty}\int_t^b f(x)\:dx
        \end{align*}
        provided this limit exists (as a finite number).
    \end{enumerate}
    The improper integrals $\int_a^\infty f(x)\:dx$ and $\int_{-\infty}^b f(x)\:dx$ are called \textbf{convergent} if the corresponding limit exists and \textbf{divergent} otherwise.\\
    If both  $\int_a^\infty f(x)\:dx$ and $\int_{-\infty}^a f(x)\:dx$ are convergent, then we define
    \begin{align*}
        \int_{-\infty}^\infty f(x)\:dx = \int_a^\infty f(x)\:dx + \int_{-\infty}^a f(x)\:dx
    \end{align*}
\end{definition}
\begin{definition}
    \textbf{of an improper integral of type 2.}
    \begin{enumerate}
        \item If $f$ is continuous on $[a,b)$ and discontinuous at $b$, then
        \begin{align*}
            \int_a^b f(x)\:dx = \lim_{t\to b^-}\int_a^t f(x)\: dx
        \end{align*}
        if this limit exists.
        \item If $f$ is continuous on $(a,b]$ and discontinuous at $a$, then
        \begin{align*}
            \int_a^b f(x)\:dx = \lim_{t\to a^+}\int_t^b f(x)\: dx
        \end{align*}
        if this limit exists.
    \end{enumerate}
    The improer integral $\int_a^b f(x)\:dx$ is called \textbf{convergent} if the corresponding limit exists and \textbf{divergent} otherwise.\\
    If $f$ has a discontinuity at $c$, where $a<c<b$, and both $\int_a^c f(x)\:dx$ and $\int_c^b f(x)\:dx$ are convergent, then we define
    \begin{align*}
        \int_a^b f(x)\: dx = \int_a^c f(x)\:dx + \int_c^b f(x)\:dx.
    \end{align*}
\end{definition}
\begin{theorem}
    \textbf{The comparison theorem.} Suppose that $f$ and $g$ are continuous functions with $f(x)\geq g(x)\geq 0$ for $x\geq a$.
    \begin{enumerate}
        \item If $\int_a^\infty f(x)\:dx$ is convergent, then $\int_a^\infty g(x)\:dx$ is convergent.
        \item If $\int_a^\infty g(x)\:dx$ is divergent, then $\int_a^\infty f(x)\:dx$ is divergent.
    \end{enumerate}
\end{theorem}
\section{Applications}
\begin{definition}
    The area $A$ of the region bounded by the curves $y=f(x)$, $y=g(x)$, and the lines $x=a$, $x=b$, where $f$ and $g$ are continuous and $f(x)\geq g(x)$ for all $x\in[a,b]$, is
    \begin{align*}
        A=\int_a^b \left(f(x)-g(x)\right)\:dx.
    \end{align*}
\end{definition}
\begin{definition}
    Let $S$ be a solid that lies between $x=a$ and $x=b$. If the cross-sectional area of $S$ in the plane $P_x$, through $x$ and perpendicular to the $x$-axis, is $A(x)$, where $A$ is an integrable function, then the \textbf{volume} of $S$ is
    \begin{align*}
        V=\int_a^b A(x)\:dx.
    \end{align*}
\end{definition}
\begin{theorem}
    \textbf{The arc length formula.} If $f'$ is continuous on $[a,b]$, then the length of the curve $y=f(x)$, $a\leq x \leq b$, is
    \begin{align*}
        L = \int_a^b \sqrt{1+\left(\frac{df}{dx}\right)^2}\:dx.
    \end{align*}
\end{theorem}
\begin{theorem}
    \textbf{of Pappus.} Let $R$ be a plane region that lies entirely on one side of a line $l$ in the plane. 
    If $R$ is rotated about $l$, then the volume of the resulting solid is the product of the area $A$ of $R$ and the distance $d$ traveled by the centroid of $R$.
\end{theorem}
\end{document}