\documentclass{article}
\usepackage{homework-preamble}

\begin{document}
\title{Linear Analysis: Homework 1}
\author{Franz Miltz (UUN: S1971811)}
\date{27 September 2022}
\maketitle

Let $C^0=C^0 \rr{ \bb{ 0,1 }, \R }$.

\begin{claim*}[2d]
  The set $S$ of all $\F$-valued sequences with limit $0$ is a vector space.
  \begin{proof}
    Let $(x_n)_{n\in\N},(y_n)_{n\in\N}\in S$. Let $\lambda\in\F$. Using limit laws we have
    \begin{align*}
      \lim_{n\to\infty}(\lambda x_n) = \lambda \lim_{n\to\infty} x_n = 0
    \end{align*}
    and
    \begin{align*}
      \lim_{n\to \infty}(x_n + y_n) = \lim_{n\to\infty} x_n + \lim_{n\to\infty} y_n = 0.
    \end{align*}
    Thus $(\lambda x_n)_{n\in\N},(x_n + y_n)_{n\in\N}\in S$.
  \end{proof}
\end{claim*}

\begin{claim*}[5]
  Let $f,g\in C^0$ be given by
  \begin{align*}
    f(x)=\cos\rr{2\pi x}, \hs g \rr{ x } = \cos \rr{ 4\pi x }.
  \end{align*}
  Then $f$ and $g$ are linearly independent.
  \begin{proof}
    Firstly we note that $f,g$ are indeed continuous on $[0,1]$ so they are elements of $C^0$.
    Consider $\lambda,\mu\in\R$ such that
    \begin{align*}
      \lambda f + \mu g = 0.
    \end{align*}
    I.e., for all $x\in[0,1]$, \begin{align*}
      \lambda f \rr{ x } + \mu g \rr{ x } = 0.
    \end{align*}
    We choose $x=1/4$ to find $\lambda = 0$. Using this and choosing $x=1/8$ we obtain $\mu=0$.
    Thus $f$ and $g$ are linearly independent.
  \end{proof}
\end{claim*}

\begin{claim*}[6]
  Let $\vabs{-}_{C^0}:C^0\to \R$ be given by
  \begin{align*}
    \vabs{f}_{C^0}=\max_{x\in \bb{ 0,1 }}\abs{f \rr{ x }}.
  \end{align*}
  $\vabs{-}_{C^0}$ defines a norm.
  \begin{proof}
    To begin we note that all $f\in C^0$ are continuous on a closed and bounded interval
    and thus attain a maximum and minimum values.

    Let $f\in C^0$ and let $\lambda\in\R$. Then
    \begin{align*}
      \vabs{\lambda f}_{C^0} &= \max_{x\in \bb{ 0,1 }}\abs{\lambda f \rr{ x }} \\
                             &= \abs{\lambda} \max_{x\in \bb{ 0,1 }}\abs{f \rr{ x }} \\
                             &= \abs\lambda \vabs{f}_{C^0}.
    \end{align*}
    Let $f,g\in C^0$. Let $x_0,y_0\in \bb{ 0,1 }$ such that
    \begin{align*}
      \vabs{f}_{C^0} = \abs{f \rr{ x_0 }}, \hs \vabs{g}_{C^0} = \abs{g \rr{ y_0 }}.
    \end{align*}
    Then
    \begin{align*}
      \vabs{f + g}_{C^0} &= \max_{x\in \bb{ 0,1 }} \abs{ f \rr{ x } + g \rr{ x } } \\
                         &\leq \max_{x\in \bb{ 0,1 }} \rr{ \abs{f \rr{ x }} + \abs{g \rr{ x } }} \\
                         &= \max_{x\in \bb{ 0,1 }} \abs{ f \rr{ x } } + \max_{x\in \bb{ 0,1 }} \abs{ g \rr{ x } } \\
                         &= \vabs{f}_{C^0} + \vabs{g}_{C^0}.
    \end{align*}
    Finally, assume $f\in C^0$ such that $\vabs{f}_{C^0}=0$. Then, for all $x\in \bb{ 0,1 }$,
    \begin{align*}
      \abs{f \rr{ x }} \leq 0.
    \end{align*}
    Immediately, $f \rr{ x }=0$ everywhere, i.e. $f=0$.
  \end{proof}
\end{claim*}

\begin{lemma} \label{uniform-iff-pointwise}
  Let $ \rr{u_n}_{n\in\N}$ a sequence of functions in $C^0$. Then $ \rr{u_n}_{n\in\N}$ converges
  uniformly iff there exists a function $u\in C^0$ such that
  \begin{align*}
    \lim_{n\to\infty} \vabs{u_n-u}_{C^0}=0.
  \end{align*}
  \begin{proof}
    Assume $ \rr{u_n}_{n\in\N}$ converges uniformly to some $u\in C^0$. Then, for all $\e > 0$,
    there exists an $N\in\N$ such that, for all $x\in \bb{ 0,1 }$ and for all $n>N$,
    \begin{align*}
      \abs{u_n \rr{ x } - u \rr{ x }} < \e.
    \end{align*}
    Now fix $\e>0$ and choose $N$ accordingly. Then, for $n>N$,
    \begin{align*}
      \vabs{u_n-u}_{C^0} = \max_{x\in \bb{ 0,1 }} \abs{u_n \rr{ x } - u \rr{ x } } < \max_{x\in \bb{0,1}} \rr{\e} < \e.
    \end{align*}
    Thus \begin{align*}
      \lim_{n\to\infty} \vabs{u_n-u}_{C^0} = 0.
    \end{align*}
    Conversely, assume for some $u\in C^0$,
    \begin{align*}
      \lim_{n\to\infty} \vabs{u_n-u}_{C^0} = 0.
    \end{align*}
    Now fix $\e>0$ and choose $N$ such that, for all $n>N$,
    \begin{align*}
      \vabs{u_n-u}_{C^0}<\e.
    \end{align*}
    Then, for all $x\in \bb{ 0,1 }$ and for all $n>N$,
    \begin{align*}
      \abs{u_n \rr{ x } - u \rr{ x }} \leq \max_{x\in \bb{ 0,1 }} \abs{u_n \rr{ x } - u \rr{ x }} = \vabs{u_n-u}_{C^0} < \e.
    \end{align*}
    Thus $ \rr{u_n}_{n\in\N}$ converges uniformly to $u$.
  \end{proof}
\end{lemma}

\begin{lemma} \label{cauchy-implies-uniform}
  Let $ \rr{u_n}_{n\in\N}$ such that, for all $\e>0$, there exists $N\in\N$ such that, for all $n,m\geq N$ and for all
  $x\in \bb{ 0,1 }$, $\abs{u_m \rr{ x } - u_n \rr{ x }} < \e$. Then $u_n\to u$ uniformly for some $u\in C^0$.
  \begin{proof}
    Assume the condition above holds. In $\R$, every Cauchy sequence converges. Thus $u_n\to u$
    pointwise for some $u: \bb{ 0,1 }\to \R$. We claim that this convergence is uniform.
    To see this, fix $\e > 0$ and choose $N\in\N$ such that, for all $m,n>N$ and all $x\in \bb{ 0,1 }$,
    \begin{align*}
      \abs{u_m \rr{ x } - u_n \rr{ x }} < \e/2.
    \end{align*}
    In particular, fix $n>N$ and calculate
    \begin{align*}
      \lim_{m\to\infty} \abs{ u_m \rr{ x } - u_n \rr{ x }} = \abs{ u \rr{ x } - u_n \rr{ x } }.
    \end{align*}
    Now
    \begin{align*}
      \lim_{m\to\infty} \abs{ u_m \rr{ x } - u_n \rr{ x }} \leq \e / 2
    \end{align*}
    so finally
    \begin{align*}
      \abs{u \rr{ x } - u_n \rr{ x }} < \e.
    \end{align*}

  \end{proof}
\end{lemma}

\begin{claim*}[7]
  $ \rr{ C^0, \vabs{-}_{C^0} }$ is a Banach space.
  \begin{proof}
    We have shown above that $\vabs{-}_{C^0}$ is a norm.

    Let $ \rr{u_n}_{n\in\N}$ be a Cauchy sequence. Now fix $\e>0$ and choose $N\in\N$
    such that, for all $m,n>N$,
    \begin{align*}
      \vabs{u_m - u_n}_{C^0}<\e.
    \end{align*}
    Then, in particular, for all $x\in \bb{ 0,1 }$, $\abs{u_m \rr{ x } - u_n \rr{ x }}<\e$.
    By~\ref{cauchy-implies-uniform}, $ \rr{u_n}_{n\in\N}$ converges uniformly to some $u\in C^0$.
    We can now apply~\ref{uniform-iff-pointwise} to find that
    \begin{align*}
      \lim_{n\to\infty} \vabs{u_n-u}_{C^0} = 0.
    \end{align*}
    Thus $u_n\to u$.
  \end{proof}
\end{claim*}

\end{document}
