\documentclass{article}
\usepackage{notes-preamble}
\usepackage{enumitem}
\begin{document}
\mkthmstwounified
\title{Linear Analysis (SEM7)}
\author{Franz Miltz}
\maketitle
\tableofcontents
\pagebreak

\section{Normed linear spaces, Banach spaces}
\label{sec:definitions}

Let $\F\in \cc{ \R,\C }$.

\begin{definition}[Normed linear space]
	\label{def:nls}
	Let $X$ be an $\F$-vector space.
	A function $ \vabs{ - } : X \to \R$ is defined to be a norm on $X$ iff the following hold
	\begin{enumerate}
		\item $\forall x\in X.\: \vabs{ x }\geq 0$ and $ \vabs{ x } = 0$ iff $x = 0$.
		\item $\forall \lambda\in\F.\:\forall x \in X.\: \vabs{ \lambda x } = \abs{ \lambda } \vabs{ x }$.
		\item $\forall x,y\in X.\: \vabs{ x + y } \leq \vabs{ x } + \vabs{ y }$
	\end{enumerate}
	A pair $ \rr{ X, \vabs{ - } }$ is a normed linear space iff $X$ is a vector space and $ \vabs{ - } $
	is a norm on $X$.
\end{definition}


\begin{theorem}
	\label{thm:symmetry}
	Let $X$ be a normed linear space. For $x,y\in X$, $ \vabs{ x - y } = \vabs{ y - x } $.
\end{theorem}


\begin{definition}
	\label{def:convergence}
	Let $X$ be a normed linear space. A sequence is defined to be a function $\N\to X$. A
	sequence $ \rr{x_n}_{n\in\N}$ is defined to converge to $x\in X$ iff, for all $\e > 0$,
	there exists $N\in\N$ such that, for all $n>N$, $ \vabs{ x_n - x } < \e$.
\end{definition}


\begin{theorem}
	\label{thm:uniquenes-of-the-limit}
	Let $X$ be a normed linear space. A sequence $ \rr{x_n}_{n\in\N}$ converges to at most one point.
\end{theorem}


\begin{definition}
	\label{def:cauchy}
	Let $X$ be a normed linear space. A sequence $ \rr{x_n}_{n\in\N}$ is defined to be Cauchy iff,
	for all $\e>0$, there exists $N\in\N$ such that, for all $m,n>\N$, $ \vabs{ x_m - x_n } < \e$.
\end{definition}


\begin{theorem}
	\label{thm:convergent-implies-cauchy}
	Let $X$ be a normed linear space. If a sequence is convergent then it is Cauchy.
\end{theorem}


\begin{definition}
	\label{def:complete}
	Let $X$ be a normed linear space. A set $E\subseteq X$ is defined to be complete iff every Cauchy
	sequence in $E$ is convergent.
\end{definition}


\begin{definition}
	\label{def:banach}
	A pair $ \rr{ X, \vabs{ - } } $ is defined to be a Banach space iff it is a complete normed linear space.
\end{definition}


\section{Inequalities}
\label{sec:ineqaulities}


\begin{theorem}
	\label{thm:gagm}
	If $A,B\geq 0$ and $0\leq\theta\leq 1$ then
	\begin{align*}
		A^\theta B^{1-\theta} \leq \theta A + \rr{ 1-\theta }B.
	\end{align*}
\end{theorem}

\begin{definition}
	\label{def:holder-conjugate-exponent}
	For $p\in \rr{ 1,\infty }$ the H\"older conjuate exponent of $p$ is defined to be $p/ \rr{ p-1 }$.
	For $p=1$, the H\"older conjugate exponent is defined to be $\infty$. For $p=\infty$, the
	H\"older conjugate exponent is defined to be $1$.
\end{definition}

\begin{theorem}
	\label{thm:hce-equals-one}
	If $p\in \rr{ 1,\infty }$, and $q$ is the H\"older conjugate exponent of $p$ then
	\begin{align*}
		1 = \frac{1}{p} + \frac{1}{q}.
	\end{align*}
\end{theorem}


\begin{theorem}
	\label{thm:holder-inequality-finite}
	Let $p\in \br{ 0,\infty }$, and let $q$ be the H\"older conjugate exponent of $p$.

	For any two finite $\F$-values sequences $ \rr{ a_1, ..., a_n }$ and $ \rr{ b_1, ..., b_n }$,
	\begin{align*}
		\abs{\sum_{i=1}^n a_i b_i}
		\leq \rr{ \sum_{i=1}^n \abs{a_j}^p }^{1/p} \rr{ \sum_{i=1}^n \abs{b_j}^p }^{1/p}
	\end{align*}
\end{theorem}

\begin{theorem}
	\label{thm:minkowski-for-p-infinite}
	For any two $\F$-values sequences $ \rr{a_n}_{n\in\N}$ and $ \rr{b_n}_{n\in\N}$,
	\begin{align*}
		\max_i \abs{ a_i + b_i } \leq \max_i \abs{ a_i } + \max_i \abs{ b_i }.
	\end{align*}
\end{theorem}

\begin{theorem}
	\label{thm:dl-norms}
	Let $p\in \bb{ 1,\infty }$.
	\begin{enumerate}
		\item Let $ \vabs{ - }_{ \ell^p } : \ell^p \to \R$ be defined by \begin{align*}
			      \vabs{ x }_{ \ell^p } = \rr{ \sum_{ n }^{  } \abs{ x_n }^p }^{1/p}.
		      \end{align*}
		      Then $ \rr{ \ell^p, \vabs{ - }_{ \ell^p } }$ is a normed linear space.
		\item Let $ \vabs{ - }_{ d_p } : \F^n \to \R$ be defined by \begin{align*}
			      \vabs{ x }_{ d_p } = \rr{ \sum_{ i=1 }^{ n } \abs{ x_i }^p }^{1/p}.
		      \end{align*}
		      Then $ \rr{ \F^n, \vabs{ - }_{ d_p } } $ is a normed linear space.
	\end{enumerate}
\end{theorem}

\section{Topology} \label{sec:topology}

\begin{definition}
	\label{def:open}
	Let $X$ be a normed linear space. Given $x\in X$ and $r>0$, the open ball with radius
	$r$ and centre $x$ is defined to be
	\begin{align*}
		B\rr{x,r}=\cc{y\in X:\vabs{y-x}<r}.
	\end{align*}
	The nunit ball is $B\rr{0,1}$.

	A subset $G\subseteq X$ is open iff for all $x\in X$ there is an $r>0$ such that
	$B\rr{x,r}\subseteq G$. A subset $F\subseteq X$ is closed iff $F^C$ is open.
\end{definition}

\begin{theorem}
	Let $X$ be a normed linear space.
	\begin{enumerate}
		\item $\emptyset$ and $X$ are open.
		\item The union of any collection of open sets is open.
		\item The intersection of a finite number of open sets is open.
	\end{enumerate}
\end{theorem}

\begin{theorem}
	Let $X$ be a normed linear space. Let $x\in X$ and $r>0$.
	\begin{enumerate}
		\item $B\rr{x,r}$ is open.
		\item $x\in B\rr{x,r}$.
		\item $\cc{y\in X : \vabs{y-x}\leq r}$ is closed.
	\end{enumerate}
\end{theorem}

\begin{theorem}
	Let $X$ be a normed linear space. Let $F\subseteq X$. $F$ is closed iff every sequence of
	elements in $F$, convergent in $X$, has its limit in $F$.
\end{theorem}

\begin{theorem}
	Let $X$ be a Banach space. Let $F\subseteq X$. $F$ is clossed iff every Cauchy sequence in
	$F$ has a limit in $F$.
\end{theorem}

\begin{theorem}
	Let $X$ be a Banach space. Let $W$ be a subspace of $X$. $W$ is a Banach space iff it is
	closed.
\end{theorem}

\begin{definition}
	\label{def:closure}
	Let $X$ be a normed linear space. Let $W\subseteq X$. The closure of $W$ is
	\begin{align*}
		\overline W = \bigcap \cc{F\subseteq X : W\subseteq F, \text{$F$ closed}}
	\end{align*}
\end{definition}

\begin{definition}
	Let $\rr{X, \vabs{-}_{X}}$ and $\rr{Y, \vabs{-}_{Y}}$ be normed linear spaces. Let $E,F\subseteq X$.

	A set $E\subseteq F$ is defined to be dense in $F$ iff, for all $f\in F$ and for all
	$\e > 0$, there exists an $e\in E$ such that $\vabs{e-f}_{X}<\e$. A set $E$ is defined
	to be dense if it is dense in $X$.

	A linear map $T:X\to Y$ is defined to be an isometry iff, for all $x\in X$,
	$\vabs{Tx}_{Y}=\vabs{x}_{X}$. $X$ and $Y$ are defined to be isometrically isomorphic iff
	there exists an onto isometry $T:X\to Y$.

	A Banach space completion of $\rr{X,\vabs{-}_{X}}$ is a pair $\rr{\rr{Y,\vabs{-}_{Y}}}$
	such that $\rr{Y,\vabs{-}_{Y}}$ is a Banach space, $T:X\to Y$ is an isometry, and $T\rr{X}$
	is dense in $Y$.
\end{definition}

\begin{theorem}[Existence of unique completion]
	Let $X$ be a normed linear space.

	There exists a Banach space completion of $X$.

	If $\rr{\rr{Y,\vabs{-}_{Y}}, T_Y}$ and $\rr{\rr{Z,\vabs{-}_{Z}}, T_Z}$ are Banach space
	completion of $X$, then $\rr{Y,\vabs{-}_{Y}}$ and $\rr{Z,\vabs{-}_{Z}}$ are isometrically
	isomorphic.
\end{theorem}

\section{Compactness}
\label{sec:compactness}

\begin{definition}
	Let $X$ be a normed linear space. Let $E\subseteq X$. A cover $E$ is defined to be a
	collection of sets $\cc{U_\alpha}_{\alpha\in A}$ such that
	$E\subseteq\bigcup_{\alpha\in A}U_\alpha$. Given a cover $\cc{U_\alpha}_{\alpha\in A}$
	is defined to be a subset $A'\subseteq A$ such that $\cc{U_\alpha}_{\alpha\in A'}$ is also
	a cover of $E$. An open cover is a cover $\cc{U_\alpha}_{\alpha\in A}$ such that each
	$U_\alpha$ is open.

	The set $E$ is compact iff every open cover of $E$ has a finite subcover.
\end{definition}

\begin{theorem}[Heine-Borel in $\R$]
	Every closed and bounded interval is compact in $\R$.
\end{theorem}

\begin{theorem}
	Let $\rr{X,\vabs{-}_{X}}$ and $\rr{Y,\vabs{-}_{Y}}$ be normed linear spaces. A function
	$f:X\to Y$ is continuous at $x_0\in X$ iff, for lal $\e > 0$, there exists a $\delta > 0$,
	such that, for all $x\in X$, $\vabs{x-x_0}_{X}<\delta$ implies
	$\vabs{f\rr{x}-f\rr{x_0}}_{Y}<\e$.
\end{theorem}

\begin{theorem}[Extreme value theorem]
	Let $X$ be a normed linear space and $f:X\to\R$ be a function. If $K\subseteq X$ is compact
	and $f$ is continuous, then $f$ has a maximum and a minimum on $K$.
\end{theorem}

\begin{theorem}
	Let $X$ be a normed linear space. Any closed subset of a compact set is compact.
\end{theorem}

\begin{definition}
	\label{def:bounded-set}
	Let $X$ be a normed linear space. Let $E\subseteq X$. $E$ is a bounded set iff there is an
	$r>0$ such that $E\subseteq B\rr{0,r}$.
\end{definition}

\begin{lemma}
	Let $X$ be a normed linear space.
	\begin{enumerate}
		\item If a Cauchy sequence has a subsequential limit, then it converges.
		\item Any subsequence of a convergent sequence converges to the same point.
	\end{enumerate}
\end{lemma}

\begin{theorem}
	Let $X$ be a normed linear space. Let $E\subseteq X$. If $E$ is compact then it is closed and bounded.
\end{theorem}

\begin{definition}
	Let $X$ be a finite-dimensional vector space. Let $\mathcal B=\rr{e_1,...,e_n}$ be
	an ordered basis for $X$.

	The $\vabs{-}_{d_\infty,\mathcal B}$ norm is given by
	\begin{align*}
		\vabs{x}_{d_\infty,\mathcal B}=\max_i \abs{x_i},
	\end{align*}
	where the $\rr{x_i}_{i=1}^n$ are the coefficients such that
	\begin{align*}
		x = \sum_{i=1}^{n} x_i e_i.
	\end{align*}
\end{definition}

\begin{lemma}[Heine-Borel in $\rr{X,\vabs{-}_{d_\infty,\mathcal B}}$]
	Let $X$ be a finite-dimensional vector space with an ordered basis $\mathcal B = \rr{e_1, ..., e_n}$.

	In $\rr{X,\vabs{-}_{d_\infty,\mathcal B}}$ there is the following equivalence: $E\subseteq X$
	is compact iff $E$ is closed and bounded.
\end{lemma}

\section{Equivalence}
\label{sec:equivalence}

\begin{definition}
	\label{def:equivalence}
	Let $X$ be an $F$-vector space. Let $\vabs{-}_{1}$ andd $\vabs{-}_{2}$ be norms on $X$.

	Then $\vabs{-}_{1}$ and $\vabs{-}_{2}$ are equivalent iff there is a constant $A>0$ such
	that, for all $x\in X$,
	\begin{align*}
		\vabs{x}_{1}\leq A\vabs{x}_{2}, \hs
		\vabs{x}_{2}\leq A\vabs{x}_{1}.
	\end{align*}
	A is called an equivalence constant.
\end{definition}

\begin{lemma}
	Let $X$ be an $F$-vector space. Let $\vabs{-}_{1}$ and $\vabs{-}_{2}$ be equivalent norms
	and $A>0$ an equivalence constant.

	For all $x\in X$ and $r>0$,
	\begin{align*}
		B^1\rr{x,r}\subseteq B^2\rr{x,Ar}, \hs
		B^2\rr{x,r}\subseteq B^1\rr{x,Ar}.
	\end{align*}
\end{lemma}

\begin{lemma}
	Let $X$ be an $F$-vector space and let $\vabs{-}_{1},\vabs{-}_{2}$ be equivalent norms
	on $X$. Then the each of the following statements is true with respect to $\vabs{-}_{1}$ iff it is true with respect to $\vabs{-}_{2}$.
	\begin{enumerate}
		\item $\rr{n}_{n\in\N}$ converges to $x$.
		\item $\rr{n}_{n\in\N}$ is Cauchy.
		\item $X$ is Banach.
		\item $G\subseteq X$ is open.
		\item $F\subseteq X$ is closed.
		\item $K\subseteq X$ is compact.
	\end{enumerate}
\end{lemma}

\begin{lemma}
	Let $X$ be a vector space. Equivalence of norms is an equivalence relation on the set of
	norms on $X$.
\end{lemma}

\begin{theorem}
	Let $X$ be a finite-dimensional $F$-vector space. All norms over $X$ are equivalent.
\end{theorem}

\begin{theorem}[Heine-Borel in finite-dimensional spaces]
	Let $\rr{X, \vabs{-}}$ be a normed linear space.

	If $X$ is finite dimensional then there is the following equivalence: $K\subseteq X$
	is compact iff it is closed and bounded.
\end{theorem}

\begin{theorem}
	Let $X$ be a normed linear space. The unit sphere is compact iff $X$ is finite dimensional.
\end{theorem}

\end{document}
