\documentclass{article}
\usepackage{notes-preamble}
\usepackage{enumitem}
\usepackage{mathrsfs}
\begin{document}
\mkthmstwounified
\title{Metric Spaces (SEM6)}
\author{Franz Miltz}
\maketitle
\noindent Textbook: V. Bryant, \emph{Metric Spaces: Iteration and Application}
\tableofcontents
\pagebreak

\section{Metrics, norms and inner products}

\subsection{Metrics and metric spaces}

\begin{definition}[Metric space]
    A \emph{metric} on a non-empty set $X$ is a function $d:X\times X\to\R$
    such that, for all $x,y,z\in X$,
    \begin{enumerate}
        \item $d(x,y)=0$ iff $x=y$,
        \item $d(x,y)=d(y,x)$,
        \item $d(x,y)\leq d(x,z) + d(z,y)$.
    \end{enumerate}
    A \emph{metric space} is an ordered pair $(X,d)$ where $X$ is a non-empty set and
    $d$ is a metric on $X$.
\end{definition}

\begin{lemma}
    The following are metric spaces:
    \begin{enumerate}
        \item $(\R^n, d_1)$ where $d_1:\R^n\times\R^n\to\R$ is defined by \begin{align*}
                  d_1(x,y)=\abs{x_1-y_1}+\cdots+\abs{x_n-y_n}.
              \end{align*}
        \item $(\R^n, d_\infty)$ where $d_\infty:\R^n\times\R^n\to\R$ is defined by \begin{align*} d_\infty(x,y)=\max\{\abs{x_1-y_1},...,\abs{x_n-y_n}\}.
              \end{align*}
    \end{enumerate}
\end{lemma}

\begin{lemma}
    Let $a,b\in\R$ with $a<b$.
    Then the set of all real valued continuous functions defined of real valued
    continuous functions defined on $[a,b]$, denoted $C([a,b])$, together with
    the usual operations of addition and scalar multiplication is a vector space.
    Further let $d:C([a,b])\times C([a,b])\to \R$ be given by
    \begin{align*}
        d(f,g)=\sup\{\abs{f(x)-g(x)}:x\in[a,b]\}.
    \end{align*}
    Then $(C([a,b]),d)$ is a metric space.
\end{lemma}

\begin{definition}[Metric subspace]
    Let $(X,d)$ be a metric space. For any non-empty subset $Y\subseteq X$ we define
    a function $d_Y:Y\times Y\to\R$ by
    \begin{align*}
        d_Y(y_1,y_2)=d(y_1,y_2).
    \end{align*}
    Then $(Y,d_Y)$ is a metric subspace of $(X,d)$.
\end{definition}

\begin{definition}[Ultrametric]
    Let $(X,d)$ be a metric space. If for all $x,y,z\in X$,
    \begin{align*}
        d(x,y)\leq \max\{d(x,z),d(z,y)\}
    \end{align*}
    then $d$ is called an \emph{ultrametric}.
\end{definition}

\begin{proposition}[Reverse triangle inequality]
    Let $(X,d)$ be a metric space and let $x,y,z\in X$. Then
    \begin{align*}
        \abs{d(x,y)-d(x,z)}\leq d(y,z).
    \end{align*}
\end{proposition}

\subsection{Norms and normed linear spaces}

\begin{definition}[Normed linear space]
    Let $X$ be a real vector space. A \emph{norm} is a function $\vabs{-}:X\to\R$
    where for all $x,y\in X$ and all $\lambda\in\R$,
    \begin{enumerate}
        \item $\vabs x \geq 0$,
        \item $\vabs x = 0$ iff $x = 0$,
        \item $\vabs{\lambda x}=\abs \lambda \vabs x$,
        \item $\vabs{x+y}\leq \vabs x +\vabs y$.
    \end{enumerate}
    A \emph{normed linear space} is a pair $(X,\vabs{-})$ where $X$ is a real vector
    space and $\vabs{-}$ is a norm on $X$.
\end{definition}

\begin{lemma}
    Let $p\geq 1$ be a real number.
    Let the set of all real sequences $(x_n)_{n\in\N}$ such that the series
    $\sum_{n=1}^\infty \abs{x_n}^p$ converges be denoted by $L^p$.
    We define the following operations:
    \begin{align}
        \label{sequence_operations}
        +:L^1\times L^1\to L^1;(x,y)\mapsto x+y               & =(x_n+y_n)_{n\in\N}=(x_1+y_1,x_2+y_2,...,x_n+y_n,...),                 \\
        \cdot:\R\times L^1\to L^1;(\lambda,x)\mapsto\lambda x & =(\lambda x_n)_{n\in\N}=(\lambda x_1,\lambda x_2,...,\lambda x_n,...).
    \end{align}
    Then $(L^p,+,\cdot)$ is a vector space.
    We define $\vabs{-}_p:L^p\to\R$ by
    \begin{align*}
        \vabs x_p = \left(\sum_{n=1}^\infty \abs{x_n}^p\right)^{1/p}.
    \end{align*}
    Then $(L^p, \vabs{-}_p)$ is a normed linear space.
    Further, we define $d_p:L^p\times L^p\to\R$ by
    \begin{align*}
        d_p(x,y)=\vabs{x-y}_p=\left(\sum_{n=1}^\infty\abs{x_n-y_n}^p\right)^{1/p}.
    \end{align*}
    Then $(L^p,d_p)$ is a metric space.
\end{lemma}

\begin{lemma}
    Let the set of all bounded real sequences be denoted by $L^\infty$. Then, together with
    the usual operations defined in (\ref{sequence_operations}), $L^\infty$ is a vector space.
    We define $\vabs{-}_\infty:L^\infty\to\R$ by
    \begin{align*}
        \vabs x_\infty = \sup\{\abs{x_n}:n\in\N\}.
    \end{align*}
    Then $(L^\infty,\vabs{-}_\infty)$ is a normed linear space.
    Further, let $d_\infty:L^\infty\times L^\infty\to\R$ is given by
    \begin{align*}
        d_\infty(x,y)=\vabs{x-y}_\infty = \sup\{\abs{x_1-y_1,x_2-y_2,...}\}.
    \end{align*}
    Then $(L^\infty, d_\infty)$ is a metric space.
\end{lemma}

\begin{lemma}[Minkowski's inequality]
    Let $p\geq 1$ be a real number. Let $x,y\in L^p$. Then
    \begin{align*}
        \left(\sum_{i=1}^\infty \abs{x_i+y_i}^p\right)^{1/p}
        \leq \left(\sum_{i=1}^\infty \abs{x_i}^p\right)^{1/p}
        + \left(\sum_{i=1}^\infty \abs{y_i}^p\right)^{1/p}.
    \end{align*}
\end{lemma}

\begin{lemma}[H\"older's inequality]
    Let $p,q\geq 1$ be real numbers with $1/p+1/q=1$. Let $x\in L^p$ and $y\in L^q$. Then
    \begin{align*}
        \sum_{n=1}^\infty \abs{x_ny_n}\leq
        \left(\sum_{n=1}^\infty\abs{x_n}^p\right)^{1/p}
        \left(\sum_{n=1}^\infty\abs{y_n}^q\right)^{1/q}.
    \end{align*}
\end{lemma}

\subsection{Inner products and inner product spaces}

\begin{definition}[Inner product space]
    Let $X$ be a real vector space over $\R$. An \emph{inner product}
    is a map $\lra{.,.}:X\times X\to\R$ where for all $x,y,z\in X$ and
    $\lambda,\mu\in\R$,
    \begin{enumerate}
        \item $\lra{x,x}\geq 0$ with $\lra{x,x}=0$ iff $x=0$,
        \item $\lra{x,y}=\lra{y,x}$,
        \item $\lra{\lambda x+\mu y,z}=\lambda\lra{x,z}+\mu{y,z}$.
    \end{enumerate}
    A \emph{inner product space} is a pair $(X, \lra{.,.})$ where $X$ is a real
    vector space and $\lra{.,.}$ is a norm on $X$.
\end{definition}

\begin{lemma}
    Let $(X,\lra{.,.})$ be an inner product space. Let $\vabs{-}:X\to\R$ be given
    by
    \begin{align*}
        \vabs x = \sqrt{\lra{x,x}}.
    \end{align*}
    Then $(X,\vabs{-})$ is a normed linear space. Further, let $d:X\times X\to\R$
    be given by
    \begin{align*}
        d(x,y)=\vabs{x-y}.
    \end{align*}
    Then $(X,d)$ is a metric space.
\end{lemma}

\section{The real line}

\subsection{Open and closed sets}

\begin{definition}
    A subset $F\subseteq\R$ is said to be \emph{closed} if it contains the limit of all
    its convergent sequences, i.e. if $(x_n)_{n\in\N}$ is any sequence with $x_n\in F$
    for all $n$, and $x_n\to x$ for some $x\in\R$, then $x\in F$.
\end{definition}

\begin{proposition}[Notes 2.2]
    The union of finitely many closed sets is a closed set.
\end{proposition}

\begin{proposition}[Notes 2.4]
    The intersection of any family of closed sets is a closed set.
\end{proposition}

\begin{definition}
    A subset $G\subseteq\R$ is said to be \emph{open} if it its complement $G^c$ is closed.
\end{definition}

\begin{proposition}[Notes 2.6]
    Let $G\subseteq\R$. The following are equivalent:
    \begin{enumerate}
        \item $G$ is open.
        \item For every point $x\in G$, there exists a positive raidus $r$, such that
              $(x-r,x+r)\subseteq G$.
    \end{enumerate}
\end{proposition}

\begin{proposition}[Notes 2.8]
    The union of any family of open sets is an open set.
\end{proposition}

\begin{proposition}[Notes 2.9]
    The intersection of finitely many open sets is an open set.
\end{proposition}

\begin{theorem}[Notes 2.10]
    A subset of the real line is open iff it is the union of open intervals.
\end{theorem}

\begin{theorem}[Notes 2.11]
    Every open subset of the real line can be written as an at most countable union
    of pairwise disjoint open intervals.
\end{theorem}

\subsection{Another look at continuity and uniform continuity}

\begin{proposition}[Notes 2.23]
    If $K\subseteq\R$ is compact and $f:K\to\R$ is continuous, then $f$ is
    uniformly continuous.
\end{proposition}

\begin{proposition}[Notes 2.24]
    A function $f:\R\to\R$ is continuous iff the inverse image of every
    open set is an open set.
\end{proposition}

\begin{definition}
    Let $A\subseteq\R$. A subset $G\subseteq A$ is said to be \emph{relatively open}
    if there exists an open subset $O\subseteq\R$ such that $G=A\cap O$.
\end{definition}

\begin{theorem}
    Let $A\subseteq\R$. A function $f:A\to\R$ is continuous iff the inverse
    image $\inv f(G)$ of every open set $G$ is a relatively open subset of $A$.
\end{theorem}

\subsection{Compact sets and the Heine-Borel theorem}

\begin{definition}
    A subset $K\subseteq\R$ is said to be\emph{compact} if every sequence of elements
    of $K$ hasa subsequence that converges to an element of $K$.
\end{definition}

\begin{theorem}[Notes 2.15]
    On the real line, a set is compact if it is closed and bounded.
\end{theorem}

\begin{theorem}[Notes 2.17]
    Let $K$ be a compact set on the real line. If $f:K\to\R$ is continuous
    then it is bounded.
\end{theorem}

\begin{theorem}[Extreme Value Theorem; Notes 2.18]
    Let $K$ be a compact set on the real line. If $f:K\to\R$ is continuous
    then $f$ has a maximum and a minimum.
\end{theorem}

\begin{definition}
    Let $E$ be a subset of $\R$. An \emph{open cover} of $E$ is a family
    $(G_i)_{i\in I}$ of open sets such that
    \begin{align*}
        E\subseteq \bigcup_{i\in I}G_i.
    \end{align*}
    A \emph{subcover} of $(G_i)_{i\in I}$ is a smaller family $(G_i)_{i\in I'}$,
    where $I'\subseteq I$, that still covers $E$.
\end{definition}

\begin{theorem}[Heine-Borel; Notes 2.20]
    Let $a,b\in\R$ with $a<b$. Every open cover of $[a,b]$ has a finite subcover.
\end{theorem}

\begin{theorem}[Notes 2.21]
    Let $K\subseteq\R$. The following are equivalent:
    \begin{enumerate}
        \item $K$ is compact.
        \item Every open cover of $K$ has a finite subcover.
    \end{enumerate}
\end{theorem}

\section{Open balls}

\begin{definition}
    Let $(X,d)$ be a metric space, $c\in X$ and $r>0$. The \emph{open ball} with center $c$
    and radius $r$, denoted by $B(c,r)$, is the set of all elements of $X$ whose distance
    from $c$ is less than $r$, i.e.
    \begin{align*}
        B(c,r) = \{x\in X : d(x,c)<r\}.
    \end{align*}
    Thet set
    \begin{align*}
        \{x\in X: d(x,c)\leq r\}
    \end{align*}
    is called the \emph{closed ball} with center $c$ and radius $r$.
\end{definition}

\begin{proposition}[Notes 3.7]
    Let $(X,d)$ be a metric space, $c\in X$ and $r>0$. For every point $x$ in the open
    ball $B(c,r)$ there exists a positive number $r'$ such that $B(x,r')\subseteq B(c,r)$.
\end{proposition}

\section{Open sets and closed sets}

Let $(X,d)$ be a metric space.

\subsection{Open sets}

\begin{definition}
    Let $A\subseteq X$. A point $a$ in $A$ is
    an \emph{interior point of $A$} if there exists $r>0$ such that $B(a,r)\subseteq A$.
\end{definition}

\begin{definition}
    Let $G\subseteq X$. Then $G$ is \emph{open} if
    every point in $G$ is an interior point of $G$, i.e. for every $x\in G$ there exists
    $r>0$ such that $B(x,r)\subseteq G$.
\end{definition}

\begin{theorem}[Notes 4.6]
    Let $(G_i)_{i\in I}$ be any family of open subsets of $X$. Then the union
    $\bigcup_{i\in I} G_i$ is open.
\end{theorem}

\begin{theorem}[Notes 4.7]
    Let $G_1,...,G_n\subset X$ be open. Then the intersection $G_1\cap\cdots\cap G_n$
    is open.
\end{theorem}

\begin{proposition}[Notes 4.9]
    A set in a metric space is open iff it can be written as a union of open balls.
\end{proposition}

\begin{definition}
    Let $(Y,d_Y)$ be a subspace of $(X,d)$ and let $G\subseteq Y$. Then $G$ is called
    \emph{relatively open} if it is open in the metric space $(Y,d_Y)$.
\end{definition}

\begin{proposition}[Notes 4.10]
    Let $(Y,d_Y)$ be a subspace of $(X,d)$ and let $G\subseteq Y$. The following
    are equivalent
    \begin{itemize}
        \item $G$ is open in the metric space $(Y,d_Y)$;
        \item there exists $O\subseteq X$ that is open in the metric space $(X,d)$ such that $G=O\cap Y$.
    \end{itemize}
\end{proposition}

\begin{proposition}[Notes, Exercise 4.11]
    Let $(Y,d_Y)$ be a subspace of $(X,d)$ such that $Y$ is open and non-empty. Then a
    subset $G\subseteq Y$ is open in the metric space $(Y,d_Y)$ iff it is open in the
    metric space $(X,d)$.
\end{proposition}

\begin{definition}
    A metric space $(X,d)$ is said to be \emph{discrete} if every subset of $X$ is open.
\end{definition}

\subsection{Closed sets}

\begin{definition}
    Let $F\subseteq X$. Then $F$ is \emph{closed} if $F^c$ is open.
\end{definition}

\begin{theorem}[Notes 4.20]
    Let $(F_i)_{i\in I}$ be a family of closed subsets of $X$. Then $\bigcap_{i\in I}F_i$
    is closed.
\end{theorem}

\begin{theorem}[Notes 4.21]
    Let $F_1,...,F_n\subseteq X$ be closed. Then $F_1\cup\cdots\cup F_n$ is closed.
\end{theorem}

\begin{definition}
    Let $(Y,d_Y)$ be a subspace of $(X,d)$ and let $F\subseteq Y$. Then $F$ is called
    \emph{relatively closed} if it is closed in the metric space $(Y,d_Y)$.
\end{definition}

\begin{proposition}
    Let $(Y,d_Y)$ be a subspace of $(X,d)$ and let $F\subseteq Y$. The following are
    equivalent
    \begin{itemize}
        \item $F$ is closed in the metric space $(Y,d_Y)$;
        \item there exists $C\subset X$ that is closed in the metric space $(X,d)$ such that
              $F=C\cap Y$.
    \end{itemize}
\end{proposition}

\begin{definition}
    A set $A\subseteq X$ is called \emph{clopen} if it is both closed and open
    in the metric space $(X,d)$.
\end{definition}

\subsection{Interior, exterior, boundary}

\begin{definition}
    Let $A\subseteq X$. A point $x\in X$ is called an \emph{interior point} of $A$ if 
    there exists an open ball $B(x,r)\subseteq A$. The set of all interior points is denoted 
    $\interior A$ and called \emph{interior} of $A$.
\end{definition}

\begin{definition}
    Let $A\subseteq X$. A point $x\in X$ is called an \emph{exterior point} of $A$ if 
    there exists an open ball $B(x,r)\subseteq A^c$. The set of all exterior points is denoted 
    $\exterior A$ and called \emph{exterior} of $A$.
\end{definition}

\begin{definition}
    Let $A\subseteq X$. A point $x\in X$ is called a \emph{boundary point} of $A$ if, for all 
    $r>0$,
    \begin{align*}
        B(x,r)\cap A\not=\emptyset\hs\text{and}\hs B(x,r)\cap A^c\not=\emptyset.
    \end{align*}
    The set of all boundary points of $A$ is denoted $\p A$ and is called the boundary of $A$.
\end{definition}

\subsection{Closure}

\begin{definition}
    Let $A\subseteq X$. The smallest closed set containing $A$, denoted $\overline A$, is called
    the \emph{closure} of $A$.
\end{definition}

\begin{proposition}[Notes 4.26]
    Let $A,B\subseteq X$. Then
    \begin{itemize}
        \item $\overline\emptyset = \emptyset$ and $\overline X = X$,
        \item $\overline{\overline A} = \overline A$,
        \item $A$ is closed iff $A=\overline A$,
        \item $A\subseteq B$ implies $\overline A\subseteq\overline B$,
        \item $\overline{A\cup B} = \overline A \cup \overline B$.
    \end{itemize}
\end{proposition}

\begin{theorem}[Lectures Week 5, Theorem 2]
    Let $x\in X$, $A\subseteq X$. Then $x\in\overline A$ iff $\dist(x,A)=0$.
\end{theorem}

\subsection{Adherent points}

\begin{definition}[Notes 4.27]
    Let $A\subseteq X$ and $x\in X$. We say that $x$ is an \emph{adherent point of $A$}
    iff, for every $r>0$, we have $B(x,r)\cap A\not=\emptyset$.
\end{definition}

\begin{proposition}[Notes 4.29]
    Let $A\subseteq X$. A point $x\in X$ is an adherent point of $A$ iff
    $x\in\overline A$.
\end{proposition}

\begin{definition}
    Let $A\subset X$ and $x\in X$.
    \begin{enumerate}
        \item We say that $x$ is an \emph{accumulation point} of $A$ if, for every $r>0$,
              \begin{align*}
                  B(x,r)\cap A\not=\{x\}.
              \end{align*}
        \item We say that $x$ is an \emph{isolated point} of $A$ if there exists $r_0>0$
              such that \begin{align*}
                  B(x,r_0)\cap A = \{x\}.
              \end{align*}
    \end{enumerate}
\end{definition}

\subsection{Dense sets}

\begin{definition}
    A subset $D\subseteq X$ is said to be \emph{dense} if, for every $x\in X$
    and every $r>0$, there exists a point $y\in D$ such that $d(x,y)<r$.
\end{definition}

\begin{proposition}[Notes 4.33]
    A subset $D\subseteq X$ is dense iff it intersects every non-empty open
    subset of $X$.
\end{proposition}

\section{Convergent sequences}

\subsection{The metric space $C([a,b])$}

\begin{theorem}
    Let $a<b$ be real numbers. Then define $X=C([a,b])$ as the set of all continuous functions 
    $f:[a,b]\to\R$ and the map $d_\infty:X\to X\to\R$ by 
    \begin{align*}
        d_\infty(f,g) = \max\{\abs{f(t)-g(t)}: t\in[a,b]\}.
    \end{align*}
    Then $(X,d_\infty)$ is a metric space.
\end{theorem}

\begin{lemma}
    A sequence $(f_n\in C[a,b])_{n\in N}$ converges to $f\in C[a,b]$ iff $(f_n)_{n\in N}$
    converges uniformly to $f$.
\end{lemma}

\section{Continuity}

\subsection{Continuous functions}

\begin{definition}
    Let $(X,d_X),(Y,d_Y)$ be metric spaces and $f:X\to Y$ a function. We say $f$ is \emph{continuous
    at a point $x_0$} in $X$ if, for every positive $\e$, there exists a positive $\delta$, such that,
    for all $x\in X$ with $d_X(x,x_0)<\delta$ we have $d_Y(f(x),f(x_0))<\e$.
\end{definition}

\begin{theorem}
    Let $(X,d_X),(Y,d_Y)$ be metric spaces, $f:X\to Y$ be a function and $x_0\in X$. Then $f$ is continuous
    at $x_0$ iff, for every open neighbourhood $G$ of $f(x_0)$, there exists an open neighbourhood $O$ of
    $x_0$, such that, for all $x\in O$, we have $f(x)\in G$.
\end{theorem}

\begin{definition}
    Let $(X,d_X),(Y,d_Y)$ be metric spaces and $f:X\to Y$ a function. We say $f$ is \emph{continuous}
    if it is continuous at every point in $X$.
\end{definition}

\begin{theorem}
    Let $(X,d_X),(Y,d_Y)$ be metric spaces and $f:X\to Y$ a function. Then $f$ is continuous iff 
    the inverse image $\inv f(G)$ of any open subset $G\subset Y$ is an open subset of $X$.
\end{theorem}

\end{document}
