\documentclass{article}
\usepackage{homework-preamble}
\mkanonthms
\begin{document}
\title{Honours Algebra: Practicing Proof 2}
\author{Franz Miltz (UUN: S1971811)}
\date{11 March 2022}
\maketitle
\newcommand{\gaussint}{\Z[i]}
We write $i=\sqrt{-1}$ and $\overline x = \overline{r+si} = r - si$.
\begin{claim*}[1]
	$\Z[i]$ is an integral domain.
	\begin{proof}
		Clearly, $\gaussint$ is a non-zero commutative ring. Let $x,y\in\gaussint$ such that
		$xy=0$. Assume $x\not=0$ and $y\not=0$. We then have $\overline{xy}=0$ and thus
		\begin{align*}
			(xy)(\overline{xy})=(xy)(\overline x\,\overline y)=(x\overline x)(y\overline y) = 0.
		\end{align*}
		However, note that for any $x=r+si\in\Z$,
		\begin{align*}
			x\overline x = r^2-s^2 \in\Z.
		\end{align*}
		Thus $x\overline x$ is a zero divisor in $\Z$ contradicting the fact that $\Z$ is an
		integral domain. Thus $x=y=0$, proving the claim.
	\end{proof}
\end{claim*}

\begin{claim*}[2]
	The group of units of $\gaussint$ has order four. In particular, $\gaussint^\times = \angles{i}$.
	\begin{proof}
		Let $u=a+bi\in\C^\times$. Then there exists $x=c+di\in\C$ such that $(a+bi)(c+di) = 1$.
		Equivalently,
		\begin{align*}
			a+bi = \frac{c-di}{c^2+d^2}.
		\end{align*}
		Therefore
		\begin{align*}
			a = \frac{c}{c^2+d^2},\hs b=-\frac{d}{c^2+d^2}.
		\end{align*}
		In particular, $a^2+b^2=1$. Note that any unit in $\gaussint$ is also a unit in $\C$.
		The integer solutions to $a^2+b^2=1$ are $(a,b)\in\{(0,1),(0,-1),(1,0),(-1,0)\}$.
		These correspond to $x\in\{i,-i,1,-1\}\supseteq\gaussint^\times$.

		We now note that
		\begin{align*}
			(-1)(-1) = 1,\hs i(-i) = 1
		\end{align*}
		and further
		\begin{align*}
			i^2 = -1,\hs i^3 = -i, \hs i^4 = 1.
		\end{align*}
		Therefore all integer solutions correspond to a unit in $\gaussint$, showing
		$\gaussint^\times=\{i,-i,1,-1\}$, and further $i$ is a generator, i.e. $\gaussint^\times=\angles{i}$.
	\end{proof}
\end{claim*}

\begin{claim*}[3]
	The map $\phi:\gaussint\to\Z_{q}$ given by
	\begin{align*}
		a + bi \mapsto [a]+[n][b]
	\end{align*}
	is a surjective ring homorphism.
	\begin{proof}
		We note that $(\Z_q,+)$ is an abelian group. Since $n$ and $q$ are coprime, it follows from
		\emph{Lagrange's theorem} that the order of $[n]$ in $(\Z_q,+)$ is $q$. Thus
		\begin{align*}
			\Z_q = \angles{[n]} = \{k[n] : k\in\Z\}.
		\end{align*}
		Noting $k[n]=[kn]$ by definition of $[n]$, we find
		\begin{align*}
			\Z_q = \{[kn] : k\in\Z\} = \{[k][n] : k\in\Z\} = \{\phi(0+ki):k\in\Z\}.
		\end{align*}
		This shows that $\phi$ is surjective.

		Let $(a+bi),(c+di)\in\C$. Then
		\begin{align*}
			\phi(a+bi) + \phi(c+di) = [a] + [n][b] + [c] + [n][c] = [a+c] + [n][b+d] = \phi((a+bi) + (c+di))
		\end{align*}
		where me made extensive use of the definition of the ring $(\Z_q,+,\times)$.
		Similarly,
		\begin{align*}
			\phi(a+bi)\phi(c+di)=([a]+[n][b])([c]+[n][d])=[ac]+[n][ad+bc]+[n][n][bd].
		\end{align*}
		We now note that $[n][n]=[n^2]=[-1]$. Using this, the operations on $\Z_q$ and the above,
		\begin{align*}
			\phi(a+bi)\phi(c+di)=[ac-bd]+[n][ad+bc]=\phi((a+bi)(c+di)).
		\end{align*}
		This shows that $\phi$ is a ring homorphism.
	\end{proof}
\end{claim*}

\begin{claim*}[4]
	\begin{align*}
		\ker\phi = I_n.
	\end{align*}
	\begin{proof}
		Assume $x\in I_n$. Then $x=(r+si)(n-i)$ for some $r,s\in\Z$. We find
		\begin{align*}
			\phi(x)=\phi(rn + s + (sn - r)i) = [rn + s] + [n][sn -r] = [rn + s + sn^2 -rn] = [s(n^2+1)] = [0].
		\end{align*}
		Thus $x\in\ker\phi$.

		Assume $x\in\ker\phi$. Then $x=r+si$ for some $r,s\in\Z$. We have
		\begin{align*}
			\phi(x) = [r] + [n][s] = [0].
		\end{align*}
		This clearly implies
		\begin{align*}
			r + sn \equiv 0 \mod q
		\end{align*}
		and equivalently
		\begin{align*}
			r                      & \equiv -sn \mod q                \\
			\Leftrightarrow rn     & \equiv -sn^2 \mod q              \\
			\Leftrightarrow rn - s & \equiv -s(n^2+1) \equiv 0 \mod q
		\end{align*}
		These imply $q|r+sn$ and $q|rn-s$. We define $a,b\in\Z$ by
		\begin{align*}
			a=\frac{rn-s}{n^2+1},\hs
			b=\frac{r+sn}{n^2+1}.
		\end{align*}
		Now we note that
		\begin{align*}
			a+bi=\frac{rn-s+(r+sn)i}{n^2+1}=\frac{(r+si)(n+i)}{n^2+1}=\frac{r+si}{n-i}.
		\end{align*}
		Thus $x=(a+bi)(n-i)\in I_n$ as desired.
	\end{proof}
\end{claim*}

\begin{claim*}[5]
	$\gaussint/I_n$ is a field if and only if $q$ is prime.
	\begin{proof}
		Firstly, note that $\phi$ is surjective so $\im\phi=\Z_q$. Secondly, by the \emph{First Isomorphism
			Theorem for Rings}, there exists a ring isomorphism
		\begin{align*}
			\frac{\gaussint}{\ker\phi}\cong\Z_q.
		\end{align*}
		Therefore $\gaussint/I_n$ is a field if and only if $\Z_q$ is.
		Finally, by \emph{Proposition 3.2.16} in the notes, we know that $\Z_q$ is a field if and only if $q$ is prime.
	\end{proof}
\end{claim*}

\end{document}