\documentclass{article}
\usepackage{homework-preamble}
\usepackage{enumitem}

\begin{document}

\begin{claim*}
	The category of sets and functions $\cat{Set}$ has all small colimits.
	In particular, for every small category $\cat J$ and every diagram
	\begin{align*}
		\cat D : \cat J \to \cat{Set}
	\end{align*}
	there exists a cone $(C, (c_j)_{j\in\cat J})$ where $C,c_j\in\cat{Set}$
	that is an initial element in the category of cones $\cat{Cone}(\cat D)$.
\end{claim*}

\begin{proof}
	Consider the following coproduct:
	\begin{align*}
		S = \coprod_{j\in\cat J_0} \cat D(j).
	\end{align*}
	This is the disjoint union of all the representations of the objects in $\cat J$
	in $\cat{Set}$. Observe that this can be constructed as follows:
	\begin{align*}
		S = \bigcup_{j\in\cat J_0}\{(j, d) : d \in \cat D(j)\}.
	\end{align*}
	Therefore
	\begin{align*}
		S \subseteq \cat J_0 \times \bigcup_{j\in J_0} \cat D(j).
	\end{align*}
	which implies $S$ is a set and thereby an object in $\cat{Set}$.
	Now define a relation $\sim$ on $S$
	such that $(i,a)\sim (j,b)$ if and only if
	\begin{enumerate}[label=R\arabic*.]
		\item there exists and arrow $\alpha:i\to j\in\cat J$ such that $\cat D(\alpha): a \mapsto b$, or
		\item there exists and arrow $\alpha:j\to i\in\cat J$ such that $\cat D(\alpha): b \mapsto a$.
	\end{enumerate}
	Note the following properties:
	\begin{itemize}
		\item \emph{reflexive}: Since $1_i:i\to i$ with $\cat D(1_i)=1_{\cat D(i)}:a\mapsto a$, we have $(i,a)\sim(i,a)$.
		\item \emph{symmetric}: Assume $(i,a)\sim(j,b)$ due to R1. Then we have $\alpha:i\to j$ with $\cat D(\alpha)(i) = j$.
		      Therefore $(j,b)\sim (i,a)$ by R2. Similarly for the R2 case.
	\end{itemize}
	We define another relation $\sim'$ on $S$ such that $x\sim'y$ if and only if
	\begin{enumerate}[label=E\arabic*.]
		\item $x\sim y$, or
		\item there exists a finite family of objects in $S$, $(m_k)_{k\leq n}$, such that \begin{itemize}
			      \item for all $k<n$, $m_k\sim m_{k+1}$, and
			      \item $x=m_1$, and
			      \item $y=m_n$.
		      \end{itemize}
	\end{enumerate}
	Note the following properties for all $x,y,z\in S$:
	\begin{itemize}
		\item \emph{reflexive}: $x\sim x$ implies $x\sim'x$.
		\item \emph{symmetric}: Assume $x\sim'y$. Then there exists a finite family $(m_k)_{k\leq n}$ with the properties above.
		      Consider the family $(l_k)_{k\leq n}$ with $l_k = m_{n+1-k}$ for all $k$. Note that since $\sim$ is reflexive,
		      we still have $l_k\sim l_{k+1}$ for all $k<n$. Futher, $y=l_1$ and $x=l_n$. Thus $y\sim'x$.
		\item \emph{transitive}: Assume $x\sim'y$ and $y\sim'z$. Then there exist families $(m_k)_{k\leq n_1}$ and $(l_k)_{k\leq n_2}$.
		      Now consider the family $(o_k)_{k\leq n+m}$ with \begin{align*}
			      o_k = \begin{cases}
				            m_k       & \text{if } k\leq n_1, \\
				            l_{n_1+k} & \text{if } k > n_1.
			            \end{cases}
		      \end{align*}
		      Note that $o_{n_1}=o_{n_1+1}=y$ and $y\sim y$ so $o_k \sim' o_{k+1}$ for all $k<n+m$. We also have $x=o_1$ and $z=o_{n+m}$ so $x\sim'z$.
	\end{itemize}
	Therefore, $\sim'$ is an equivalence relation on $S$. Consider the set of equivalence classes $C=S/\sim'$.
	For each $i\in\cat J_0$ we have an arrow in $\cat{Set}$, i.e. a function,
	\begin{align*}
		c_i : \cat D(i) \to C \\
		x \mapsto [x]
	\end{align*}
	where $[x]$ is the equivalence class of $x$ under $\sim'$.
	Note that  for all $\alpha:i\to j\in\cat J$, $x\in \cat D(i)$ we have $(i,x)\sim'(j,\cat D(\alpha)(x))$ and thus
	\begin{align*}
		c_j(\cat D(\alpha)(x)) = [(i,x)] = c_i(x)
	\end{align*}
	which makes $(C, (c_i)_{i\in\cat J_0})$ a cocone.
\end{proof}

\end{document}