\documentclass{article}
\usepackage{notes-preamble}
\usepackage{enumitem}
\usepackage{tikz-cd}
\usepackage{quiver}

\begin{document}
\mkthmstwounified
\title{Categories and Quantum Informatics (SEM8)}
\author{Franz Miltz}

\maketitle

\tableofcontents

\pagebreak

\section{Linear structure}

Let $(\cat{C}, \otimes, I)$ be a monoidal category.

\subsection{Scalars}

\begin{definition}
  A \emph{scalar} in $\cat{C}$ is a morphism $I\to I$.
\end{definition}

\begin{lemma}
  Scalars in $\cat{C}$ commute. That is, for scalars $s,t:I\to I$, $s\circ t = t\circ s$.
\end{lemma}

\begin{definition}
  Let $s:I\to I$ be a scalar and let $f:X\to Y$ be a morphism. Then the \emph{left scalar
  multiplication of $f$ by $s$} is the composite
  \begin{equation}
    % https://q.uiver.app/?q=WzAsNCxbMCwwLCJYIl0sWzIsMCwiWSJdLFswLDEsIklcXG90aW1lcyBYIl0sWzIsMSwiSVxcb3RpbWVzIFkiXSxbMiwzLCJzXFxvdGltZXMgZiIsMl0sWzAsMiwiXFxsYW1iZGFeey0xfSIsMl0sWzMsMSwiXFxyaG9eey0xfSIsMl0sWzAsMSwic1xcYnVsbGV0IGYiXV0=
    \begin{tikzcd}
      X && Y \\
      {I\otimes X} && {I\otimes Y}
      \arrow["{s\otimes f}"', from=2-1, to=2-3]
      \arrow["{\lambda^{-1}}"', from=1-1, to=2-1]
      \arrow["{\rho^{-1}}"', from=2-3, to=1-3]
      \arrow["{s\bullet f}", from=1-1, to=1-3]
    \end{tikzcd}
  \end{equation}
\end{definition}

\begin{lemma}
  Let $s,t:I\to I$ be scalars and $f:X\to Y$ and $g:Y\to Z$ be morphisms. Then
  \begin{enumerate}
    \item $\identity_Y\bullet f = f$;
    \item $s\bullet t = s\circ t$;
    \item $(s\bullet t)\bullet f = s\bullet (t\bullet f)$;
    \item $(t\bullet g)\circ(s\bullet f) = (t\circ s)\bullet (g\circ f)$.
  \end{enumerate}
\end{lemma}

\subsection{Superposition}

\begin{definition}
  A \emph{superposition rule} for $\cat{C}$ consists of
  \begin{enumerate}
    \item for all $f,g:X\to Y$, an arrow $f+g:X\to Y$;
    \item for all $X,Y\in\cat{C}$, an arrow $u_{X,Y}:X\to Y$
  \end{enumerate}
  such that, for suitable morphisms,
  \begin{enumerate}
    \item $f+g=g+f$;
    \item $f+(g+h)=(f+g)+h$;
    \item $f+u = f$;
    \item $(g'\circ g)+f=(g'\circ f)+(g\circ f)$;
    \item $g+(f'\circ f) = (g\circ f')+(g\circ f)$;
    \item $u\circ f = u$ and $f\circ u = u$.
  \end{enumerate}
\end{definition}

\begin{lemma}
  In a category with a zero object and a superposition rule, $u_{X,Y}=0_{X,Y}$ for
  all $X,Y\in\cat{C}$.
\end{lemma}

\begin{lemma}
  If a monoidal category has a zero object and a superposition rule then its scalars
  form a commutative semiring with absorbing zero.
\end{lemma}

\subsection{Biproducts}

\begin{definition}
  In a category with a superposition rule and a zero bject, the biproduct
  of objects $X_1,X_2$ is an object $X_1\oplus X_2$ together with projections
  $\pi_j : X_1\oplus X_2\to X_j$ and injections $\iota_j :X_j\to X_1\oplus X_2$
  such that
  \begin{enumerate}
    \item $\pi_j\circ\iota_j = \identity_{X_j}$;
    \item $\pi_j\circ\iota_{k} = 0_{A_k,A_j}$ for $j\neq k$;
    \item $\pi_1\circ\iota_1 + \pi_2\circ\iota_1 = \identity_{X_1\oplus X_2}$.
  \end{enumerate}
\end{definition}

\begin{lemma}
  Biproducts are products and coproducts.
\end{lemma}

\begin{lemma}
  If a category has biproducts then it has a unique superposition rule.
\end{lemma}

\begin{lemma}
  Let $\cat{C}$ be a category with biproducts and a zero object. If $F:\cat{C}\to\cat{D}$
  preserves zero objects then it preserves biproducts iff it is linear.
\end{lemma}

\subsection{Matrices}

\begin{definition}
  Let $X_1,\ldots,X_M$ and $Y_1,\ldots,Y_N$ be objects and $f_{m,n}:X_m\to Y_n$
  be arrows. Then we have the \emph{matrix}
  \begin{align*}
    (f_{j,k}) =
    \begin{pmatrix}
      f_{1,1} &\cdots & f_{M,1} \\
      \vdots &\ddots & \vdots \\
      f_{1,N} & \cdots & f_{M,N}
    \end{pmatrix}
    = \sum_{m,n} \iota_n \circ f_{m,n} \circ \pi_m.
  \end{align*}
\end{definition}

\begin{lemma}
  In a category with biproducts, every morphism $\oplus_{m=1}^M X_m \to \oplus_{n=1}^N Y_n$
  has a matrix representation.
\end{lemma}

\begin{corollary}
  Morphisms between biproducts are equal iff their matrix entries are equal.
\end{corollary}

\subsection{Daggers}

\begin{definition}
  A \emph{dagger} on a category $\cat{C}$ is an involutive contravariant functor
  $\dagger : \catop{C}\to\cat{C}$ that is the identity on objects. A \emph{dagger category}
  is a category equipped with a dagger.
\end{definition}

\begin{example}
  We have the following dagger structures:
  \begin{itemize}
    \item In $\Hilb$, the dagger is given by taking adjoints.
    \item In $\Rel$, the dagger is given by $t R^\dagger s$ iff $s R t$.
  \end{itemize}
\end{example}

\begin{definition}
  A \emph{monoidal dagger category} is a monoidal category $(C,\otimes)$ equipped with
  a dagger such that $(f\otimes g)^\dagger = f^\dagger \otimes g^\dagger$.
\end{definition}

\begin{definition}
  In a dagger category, a \emph{dagger biproduct} of objects $X_1$ and $X_2$ is a biproduct
  such that $\iota_j^\dagger = \pi_j$.
\end{definition}

\begin{lemma}
  In a dagger category with dagger biproducts, the dagger of a matrix is its conjugate transpose.
\end{lemma}

\begin{definition}
  A morphism $f:X\to Y$ in a dagger category is
  \begin{itemize}
    \item \emph{idempotent} when $f=f^\dagger$;
    \item a \emph{projection} when it is idempotent and self-adjoint;
    \item \emph{unitary} when $f^\dagger \circ f = \identity$ and $f\circ f^\dagger=\identity$;
    \item an \emph{isometry} when $f^\dagger \circ f=\identity$;
    \item a \emph{partial isometry} when $f^\dagger\circ f$ is a projection;
    \item \emph{positive} when $f=g^\dagger\circ g$ for some $g$.
  \end{itemize}
\end{definition}

\subsection{Measurement}

\begin{definition}
  Consider a state $a : I\to X$ and an effect $x : X\to I$ in a monoidal dagger category.
  Define
  \begin{align*}
    \Pr(x,a) = a^\dagger \circ x^\dagger \circ a\circ x.
  \end{align*}
\end{definition}

\begin{definition}
  In a dagger category with a zero object, an isometry $k:K\to X$ is \emph{a dagger kernel}
  of $f:X\to Y$ when $f\circ k=0$ and $k$ is universal with this property.
\end{definition}

\begin{lemma}
  In a dagger category with a zero object, isometries always have a dagger kernel given
  by the zero morphism.
\end{lemma}

\begin{lemma}
  In a dagger category with a zero object and dagger kernels of arbitrary morphisms,
  for all $f:X\to Y$, $f^\dagger \circ f = 0$ implies $f=0$.
\end{lemma}

\begin{definition}
  A set of effects $\cc{x_j}$ with $x_j : X\to I$ is
  \begin{enumerate}
    \item \emph{complete} if, for all non-zero $f:W\to X$, there is some $j$
      such that $x_j \circ f \neq 0$;
    \item \emph{disjoint} if, for all $i\neq j$, $x_i\circ x_i^\dagger = \identity$
      and $x_i\circ x_j^\dagger = 0$.
  \end{enumerate}
\end{definition}

\begin{example}
  We have the following complete disjoint sets of effects:
  \begin{itemize}
    \item In \Hilb, complete disjoint sets of effects correspond to unitaries $H\simeq\C^n$,
      i.e. choices of orthonormal bases.
    \item In \Rel, a complete disjoint set of effects for a set corresponds to a partition.
  \end{itemize}
\end{example}

\begin{lemma}
  Finite complete disjoint sets of effects $x_j:X\to I$ correspond exactly to morphisms
  $x:X\to\bigoplus_j I$ for which $x^\dagger$ is an isometry and that have zero kernel.
\end{lemma}

\begin{proposition}[Born rule]
  Let $x_j:X\to I$ be a finite complete set of effects in a monoidal dagger category with
  dagger biproducts, and let $a:I\to X$ be an isometric state. Then
  $\sum_j \Pr{x_j,a}=1$.
\end{proposition}


\end{document}
