\documentclass{article}
\usepackage{homework-preamble}

\begin{document}
\title{Honours Analysis: Homework 3}
\author{Franz Miltz (UNN: S1971811)}
\date{14 October 2021}
\maketitle

\section{Workshop 3 - Question 1}

\begin{claim}
   There exists a series $\sum_{n=1}^\infty a_n$ that diverges even though the 
   series $(a_n)$ converges.
\end{claim}
\begin{proof}
   Let $a_n=1/n$. Then $a_n\to 0$ as $n\to\infty$ but the harmonic series 
   $\sum_{n=1}^\infty a_n$ is known to diverge.
\end{proof}

\begin{claim}
   There exists a series $\sum_{n=1}^\infty a_n$ that diverges even though
   $a_n\to 0$ as $n\to\infty$.
\end{claim}
\begin{proof}
   Let $a_n=1/n$. Then $a_n\to 0$ as $n\to\infty$ but the harmonic series 
   $\sum_{n=1}^\infty a_n$ is known to diverge.
\end{proof}

\begin{claim}
   There exists a series $\sum_{n=1}^\infty a_n$ that converges even though
   the series $\sum_{n=1}^\infty \abs{a_n}$ diverges.
\end{claim}
\begin{proof}
   Let $a_n=(-1)^n/n$. By the \emph{Alternating Series Test}, the series $\sum_{n=1}^\infty a_n$
   converges. However, 
   \begin{align}
      \sum_{n=1}^\infty \abs{a_n} = \sum_{n=1}^\infty \frac{1}{n}
   \end{align}
   which is the harmonic series we know to diverge.
\end{proof}

\section{Workshop 3 - Question 5}

\begin{claim}
   Suppose that a series $\sum_{n=1}^\infty a_n$ converges conditionally. Prove
   that $\sum_{n=1}^\infty n^pa_n$ diverges for all $p> 1$.
\end{claim}
\begin{proof}
   By contradiction. Assume that $\sum_{n=1}^\infty n^pa_n$ converges. Then the sequence $(n^pa_n)$
   must converge to zero and therefore it must be bounded. Choose $M\in\R$ such 
   that
   \begin{align*}
      \abs{n^pa_n}\leq M.
   \end{align*}
   Observe that $n^p > 0$ for all $n\in\N$. Therefore, the above implies 
   \begin{align*}
      n^p\abs{a_n} \leq M
   \end{align*}
   and thereby
   \begin{align}
      \label{ineq1}
      \abs{a_n} \leq \frac{M}{n^p} .
   \end{align}
   Note that the infinite sum
   \begin{align*}
      \sum_{n=1}^\infty \frac{M}{n^p} = M\sum_{n=1}^\infty \frac{1}{n^p}
   \end{align*}
   is a constant multiple of a $p$-series with $p>1$. Therefore it converges.
   Now note that, due to (\ref{ineq1}), we can apply the \emph{Comparison Test}
   to find that the series $\sum_{n=1}^\infty a_n$ converges absolutely.
   This contradicts the premise and therefore our assumption must be false.
\end{proof}

\section{Workshop 3 - Question 6}

\begin{claim}
   Let $f(x)=x^2$ when $x$ is rational and $f(x)=0$ when $x$ is irrational.
   Then $f$ is continuous and differentiable only at $x=0$.
\end{claim}
\begin{proof}
   Consider a sequence $(x_n)$ such that $x_n\to 0$ as $n\to \infty$. Let $\e>0$.
   Pick $N$ such that $\abs{x_n}<\sqrt{\e}$ for all $n>N$. Then, for all $n>N$ we have
   \begin{align*}
      \abs{f(x_n) - f(0)} = \begin{cases}
         \abs{x_n^2} &\text{if $x_n$ is rational},\\
         0           &\text{if $x_n$ is irrational}.
      \end{cases}
   \end{align*}
   Note that either way, $\abs{f(x_n)-f(0)}<\e$ and thus $f(x_n)\to 0$ as $n\to\infty$.
   This shows that $f$ is continuous at $0$. Now consider
   \begin{align*}
      \lim_{h\to 0} \frac{f(h) - f(0)}{h} = \lim_{h\to 0} \frac{f(h)}{h}. 
   \end{align*}
   Let $\e>0$. Let $\delta = \sqrt{h}$. Then for all $0 < \abs{h} < \delta$, 
   \begin{align}
      \abs{\frac{f(h)}{h}} = \begin{cases}
         h^2/h = h, &\text{if $h$ is rational},\\
         0        , &\text{if $h$ is irrational}. 
      \end{cases}
   \end{align}
   In both cases, $\abs{f(h)/h}<\e$. Thus $f$ is differentiable at $0$ and $f'(0)=0$.

   Consider a sequence of rational numbers $(x_n)$ such that $x_n\to L$ where $L\not=0$.
   Then clearly 
   \begin{align}
      \lim_{n\to\infty} f(x_n) = L^2.
   \end{align}
   However, for a sequence of irrational numbers $(y_n)$ where $y_n\to L$,
   \begin{align}
      \lim_{n\to\infty} f(y_n) = 0.
   \end{align}
   Since these limits disagree for all $L\not=0$, $f$ is only continuous at $0$.
   Note that such sequences exist due to the completeness of the rationals and
   irrationals.
\end{proof}

\end{document}