\documentclass{article}
\usepackage{homework-preamble}

\begin{document}
\title{Honours Analysis: Homework 9}
\author{Franz Miltz (UNN: S1971811)}
\date{25 November 2021}
\maketitle

\section*{Workshop 9 - Question 2}

\begin{claim*}
   Let $f:[a,b]\to\R$. If $f$ is Riemann integrable, then so is $\abs f$.
\end{claim*}
\begin{proof}
   Let $\e>0$. Assume $f$ is Riemann integrable. Then there exists a partition
   $a=x_0<\cdots<x_n=b$ such that 
   \begin{align*}
      \sum_{j=1}^n \sup_{x,y\in I_j} \abs{f(x) - f(y)} \lambda(I_j) < \e,\hs 
      \text{where }I_j = (x_{j-1}, x_j).
   \end{align*}
   Observe for all $x,y\in[a,b]$,
   \begin{align*}
      \abs{f(x)-f(y)} \geq \abs{\abs{f(x)} - \abs{f(y)}}.
   \end{align*}
   Thus 
   \begin{align*}
      \sum_{j=1}^n \sup_{x,y\in I_j} \abs{\abs{f(x)} - \abs{f(y)}} \lambda(I_j)
      \leq \sum_{j=1}^n \sup_{x,y\in I_j} \abs{f(x) - f(y)} \lambda(I_j) < \e.
   \end{align*}
   Therefore $\abs f$ is Riemann integrable.
\end{proof}

\begin{claim*}
   There exists $f:[a,b]\to\R$ such that $\abs f$ is integrable but $f$ is not.
\end{claim*}
\begin{proof}
   Consider $f = 2\chi_{\Q\cap[0,1]}-1$. Observe that $\abs{f(x)} = 1$ for all $x\in[0,1]$.
   Therefore $\abs f$ is Riemann integrable. However, $f$ itself cannot be Riemann integrable 
   as $\chi_{\Q\cap[0,1]} = (f+1)/2$ and we know that the Dirichlet function is not Riemann
   integrable.
\end{proof}

\section*{Workshop 9 - Question 3 \& 6}

\begin{claim*}
   Let $-\infty \leq a < b \leq \infty$. Suppose that $f:(a,b)\to\R$ is integrable
   on $(u,b)$ for all $u\in(a,b)$. Then $f$ is integrable on $(a,b)$ if and only if 
   there exists $M<\infty$ such that 
   \begin{align}
      \label{cond}
      \int_u^a \abs f \leq M, \hs \forall u \in (a,b).
   \end{align}
   Additionally, if this holds then 
   \begin{align}
      \label{intval}
      \int_a^b f = \lim_{u\to b^-}\int_u^b f.
   \end{align}
\end{claim*}
\begin{proof}
   ($\Rightarrow$) Assume $f$ is integrable on $(a,b)$. Immediately, $\abs f$ must be integrable
   on $(a,b)$ as well. By \emph{Notes, Theorem 4.8 (c)} we then have 
   \begin{align*}
      \int_u^b \abs f \leq \int_a^b \abs f = M,
   \end{align*}
   for all $u\in(a,b)$ which is precisely (\ref{cond}). 

   ($\Leftarrow$) Assume (\ref{cond}) holds for some $M<\infty$. Consider a sequence of numbers 
   $b>u_1>u_2>\cdots>a$ such that $u_n\to a$ as $n\to\infty$. Let $u_0=b$. 
   By \emph{Notes, Theorem 4.8 (d)}, for all $n$ we have 
   \begin{align*}
      \sum_{j=1}^n \int_{u_j}^{u_{j-1}} \abs f = \int_{u_n}^b \abs f.
   \end{align*}
   Using \emph{Notes, Theorem 4.8 (c)}, we then observe that (\ref{cond}) implies
   \begin{align}
      \label{newcond}
      \sum_{j=1}^n \int_{u_j}^{u_{j-1}} \abs f \leq M, \hs \text{for }n=1,2,...
   \end{align}
   Further, since the above holds, we also have 
   \begin{align*}
      \sum_{j=1}^\infty \int_{u_j}^{u_{j-1}} \abs f \leq M, 
   \end{align*}
   and hence by \emph{Notes, Theorem 4.8 (d)} the function $f$ is integrable on
   the union of intervals $(u_j, u_{j-1})$ which is $(a,b)$. By the same theorem, 
   we also get (\ref{intval}).
\end{proof}

\section*{Workshop 9 - Question 10}

\begin{claim*}
   The function $f(x)=(-1)^{\floor{x}} / \floor{x}$ for $x\geq 1$ is not integrable
   on $[1, \infty)$.
\end{claim*}
\begin{proof}
   Consider the sequence $(v_n)$ given by $v_n = n$. Observe that in order for
   $f$ to be integrable on $[1,\infty)$, we require 
   \begin{align*}
      \sum_{j=1}^n \int_{v_j}^{v_{j+1}} \abs f \leq M, \hs \text{for }n = 1,2,...
   \end{align*}
   for some $M<\infty$ as proven on the worksheet. However, for all $j$,
   \begin{align*}
      \sum_{j=1}^n \int_{v_j}^{v_{j+1}} \abs f 
      = \sum_{j=1}^n \frac{1}{j} 
   \end{align*}
   which is clearly unbounded as $n\to\infty$.
\end{proof}

\emph{Note: So the problem with the given function is that we require the integrals of the 
absolute values to be bounded which is not the case. The given limit exists due to the alternating 
signs.}

\end{document}