\documentclass{article}
\usepackage{homework-preamble}
\mkanonthms
\begin{document}
\title{General Topology: Assignment 5}
\author{Franz Miltz (UUN: S1971811)}
\date{27 October 2022}
\maketitle

\begin{enumerate}
  \item The topic I enjoyed the most so far is the ordering of topologies and looking at
    it from a more general point of view. Having an initial and a terminal topology for each
    topological space makes a lot of sense. This also makese the category of topological
    spaces interesting, though I have yet to be formally introduced.
  \item I found the homeworks the most challenging bit of the course. Jokes aside, dealing with
    open and closed sets without any form of analysis is unintuitive to say the least.
    In Metric Spaces we had unintuitive metrics but at least they always led us back to the
    reals. Being closed meant that sequences converge, not that a set is defined to be its
    own closure by some operator.
  \item I'm enjoying the lectures and the overall style of the course. Though, admittedly,
    this is probably due to the fact that its different to everything else. I hope to get
    answers to some of the following questions: \begin{itemize}
      \item What are Hausdorff spaces?
      \item What makes topology interesting to a working mathematician?
      \item What are open questions in topology?
      \item What overlaps does topology have with other fields?
    \end{itemize}
  \item I like your sneakers.
\end{enumerate}

\end{document}
